---
title: "project_ADC_mega_main"
---


```{r,include=F}
rm(list = ls())
library(readxl)
library(tidyverse)
library(ggseg)
library(compareGroups)
library(ggpubr)
library(compareGroups)
library(sjPlot)
library(naniar)
library(bruceR)
library(mice)
library(psych)
library(dplyr)
library(broom)
library(ggplot2)
library(scales)   # for pvalue() (or use your own formatter)
```


# Data loading
```{r,include=FALSE}
path_data = "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/data"
Freesurfer_output_718 = read_xlsx(str_c(path_data,"/data_request/718_Freesurfer_output_20250619.xlsx"))
data_718_Baseline = read_xlsx(str_c(path_data,"/data_request/718_Updated_data_20250602_V2.xlsx"),sheet = "Baseline") # 9749
data_718_Diagnosis = read_xlsx(str_c(path_data,"/data_request/718_Updated_data_20250602_V2.xlsx"),sheet = "Diagnosis") # 9749
data_718_MRI = read_xlsx(str_c(path_data,"/data_request/718_Updated_data_20250602_V3.xlsx"),sheet = "MRI",col_types = c("numeric","date","text","text","text","text","text","text","text","text"))

data_718_CSF = read_xlsx(str_c(path_data,"/data_request/718_Updated_data_20250602_V2.xlsx"),sheet = "CSF")
data_718_PET = read_xlsx(str_c(path_data,"/data_request/718_Updated_data_20250602_V2.xlsx"),sheet = "PET")
data_718_NPO = read_xlsx(str_c(path_data,"/data_request/718_Updated_data_20250602_V2.xlsx"),sheet = "NPO")
data_718_NPI = read_xlsx(str_c(path_data,"/data_request/718_Updated_data_20250602_V2.xlsx"),sheet = "NPI")
data_718_MMSE = read_xlsx(str_c(path_data,"/data_request/718_Updated_data_20250602_V2.xlsx"),sheet = "MMSE")

data_718_Diagnosis$D_dat = str_split(data_718_Diagnosis$D_dat," ",simplify = TRUE)[,1]

data_718_MRI$M_dat = str_split(data_718_MRI$M_dat," ",simplify = TRUE)[,1]
data_718_MRI$M_age = as.numeric(data_718_MRI$M_age)

data_718_CSF$L_DATEAFN = str_split(data_718_CSF$L_DATEAFN," ",simplify = TRUE)[,1]
data_718_CSF$L_age = as.numeric(data_718_CSF$L_age)

data_718_PET$PT_date = str_split(data_718_PET$PT_date," ",simplify = TRUE)[,1]
data_718_NPO$N_date = str_split(data_718_NPO$N_date," ",simplify = TRUE)[,1]
data_718_NPO$N_age = as.numeric(data_718_NPO$N_age)

data_718_MMSE$V_dat = str_split(data_718_MMSE$V_dat," ",simplify = TRUE)[,1]
```



# Functions for analysis
```{r}
analysis_glm_ATN <- function(data,column_predictor){
  
  list=c("bankssts","caudal anterior cingulate","caudal middle frontal",
      "cuneus","entorhinal","frontal pole","fusiform","inferior parietal",
      "inferior temporal","insula","isthmus cingulate","lateral occipital",
      "lateral orbitofrontal","lingual","medial orbitofrontal","middle temporal",
      "paracentral","parahippocampal","pars opercularis","pars orbitalis",
      "pars triangularis","pericalcarine","postcentral","posterior cingulate",
      "precentral","precuneus","rostral anterior cingulate","rostral middle frontal",
      "superior frontal","superior parietal","superior temporal","supramarginal",
      "temporal pole","transverse temporal")
  list2 = str_replace_all(list," ","")
  CT_names <- c(str_c("lh_",list2,"_thickness"),str_c("rh_",list2,"_thickness"))
  DTK_Labes_For_Plot = list
  plot_data_L = as.data.frame(DTK_Labes_For_Plot)
  colnames(plot_data_L)[1] = "region"
  plot_data_L$hemi="left"
  plot_data_R = as.data.frame(DTK_Labes_For_Plot)
  colnames(plot_data_R)[1] = "region"
  plot_data_R$hemi="right"
  plot_data = rbind(plot_data_L,plot_data_R)
  plot_data$area_label = CT_names
  plot_data$Effect=NA
  plot_data$P_value=NA
  attach(plot_data)
  
  for (im in 1:68) {
    Region_AV1451_v1 = plot_data$area_label[im]
    funtion_glm = paste(Region_AV1451_v1,"~",column_predictor,"+amyloid_age+Sex+Edu")
    results = glm(funtion_glm,family = gaussian,data =data)
    x = summary(results)
    P_values_GLM = x$coefficients[2,3:4]
    plot_data$Effect[im] = x$coefficients[2,1]
    plot_data$P_value[im] = x$coefficients[2,4]
  }
  
  plot_data <- plot_data %>% 
    mutate(highlight = ifelse(P_value < 0.05, "sig", "non_sig")) %>%
    mutate(p_adjust = p.adjust(P_value, "fdr", n = length(P_value))) %>%
    mutate(highlight_sig = ifelse(p_adjust < 0.05, "sig", "non_sig"))
  
  return(plot_data)
}

analysis_glm_ATN_volume <- function(data,column_predictor){
  
  list=c("bankssts","caudal anterior cingulate","caudal middle frontal",
      "cuneus","entorhinal","frontal pole","fusiform","inferior parietal",
      "inferior temporal","insula","isthmus cingulate","lateral occipital",
      "lateral orbitofrontal","lingual","medial orbitofrontal","middle temporal",
      "paracentral","parahippocampal","pars opercularis","pars orbitalis",
      "pars triangularis","pericalcarine","postcentral","posterior cingulate",
      "precentral","precuneus","rostral anterior cingulate","rostral middle frontal",
      "superior frontal","superior parietal","superior temporal","supramarginal",
      "temporal pole","transverse temporal")
  list2 = str_replace_all(list," ","")
  CT_names <- c(str_c("lh_",list2,"_volume"),str_c("rh_",list2,"_volume"))
  DTK_Labes_For_Plot = list
  plot_data_L = as.data.frame(DTK_Labes_For_Plot)
  colnames(plot_data_L)[1] = "region"
  plot_data_L$hemi="left"
  plot_data_R = as.data.frame(DTK_Labes_For_Plot)
  colnames(plot_data_R)[1] = "region"
  plot_data_R$hemi="right"
  plot_data = rbind(plot_data_L,plot_data_R)
  plot_data$area_label = CT_names
  plot_data$Effect=NA
  plot_data$P_value=NA
  attach(plot_data)
  
  for (im in 1:68) {
    Region_AV1451_v1 = plot_data$area_label[im]
    funtion_glm = paste(Region_AV1451_v1,"~",column_predictor,"+amyloid_age+Sex+Edu")
    results = glm(funtion_glm,family = gaussian,data =data)
    x = summary(results)
    P_values_GLM = x$coefficients[2,3:4]
    plot_data$Effect[im] = x$coefficients[2,1]
    plot_data$P_value[im] = x$coefficients[2,4]
  }
  
  plot_data <- plot_data %>% 
    mutate(highlight = ifelse(P_value < 0.05, "sig", "non_sig")) %>%
    mutate(p_adjust = p.adjust(P_value, "fdr", n = length(P_value))) %>%
    mutate(highlight_sig = ifelse(p_adjust < 0.05, "sig", "non_sig"))
  
  return(plot_data)
}

analysis_glm_ATN_sub <- function(data,column_predictor){
  
    list_regions=c("lh_Thalamus_volume","lh_Caudate_volume","lh_Putamen_volume","lh_Pallidum_volume","lh_Hippocampus_volume","lh_Amygdala_volume",
              "lh_Accumbens_area_volume","lh_choroid_plexus_volume",
              "lh_Thalamus_volume","lh_Caudate_volume","lh_Putamen_volume","lh_Pallidum_volume","lh_Hippocampus_volume","lh_Amygdala_volume",
              "lh_Accumbens_area_volume","lh_choroid_plexus_volume")
    list_regions_plots <- c("thalamus proper","caudate","putamen","pallidum","hippocampus","amygdala","accumbens","choroid plexus",
                            "thalamus proper","caudate","putamen","pallidum","hippocampus","amygdala","accumbens","choroid plexus")
    
    list_regions_labels <- c("Thalamus","Caudate nucleus","Putamen","Pallidum","Hippocampus","Amygdala","Nucleus Accumbens","Choroid plexus",
                            "Thalamus","Caudate nucleus","Putamen","Pallidum","Hippocampus","Amygdala","Nucleus Accumbens","Choroid plexus")
    
    plot_data = as.data.frame(list_regions)
    colnames(plot_data)[1] = "region"
    plot_data$hemi[1:8]="left"
    plot_data$hemi[9:16]="right"
    plot_data$area_label = plot_data$region
    plot_data$region = list_regions_plots
    plot_data$region_label = list_regions_labels
    plot_data$Effect=NA
    plot_data$P_value=NA
    attach(plot_data)
    
    PRS_cell_type <- column_predictor
    
    for (im in 1:16) {
      Region_AV1451_v1 = plot_data$area_label[im]
      funtion_glm = paste(Region_AV1451_v1,"~",PRS_cell_type,"+amyloid_DATEAFN+Sex+Edu+eTIV")
      results = glm(funtion_glm,family = gaussian,data =data)
      x = summary(results)
      CI <- confint(results)
      P_values_GLM = x$coefficients[2,3:4]
      plot_data$Effect[im] = x$coefficients[2,1]
      plot_data$P_value[im] = x$coefficients[2,4]
      plot_data$ci_lower[im] = CI[2,1]
      plot_data$ci_upper[im] = CI[2,2]
    }
    
    plot_data <- plot_data %>% 
      mutate(highlight = ifelse(P_value < 0.05, "sig", "non_sig")) %>%
      mutate(p_adjust = p.adjust(P_value, "fdr", n = length(P_value))) %>%
      mutate(highlight_sig = ifelse(p_adjust < 0.05, "sig", "non_sig"))
    
    return(plot_data)
}

create_atrophy_maps = function(data1,data2){
  
  colors <- c("#EB482F", "#F3E66C","white", "#7CCEDE", "#2D4CA0")
  
  plot1 <- ggseg(.data = data1,atlas = "dk",position="stacked",mapping = aes(fill = Effect, colour = highlight,size = highlight),
                 show.legend = T) +
          scale_fill_gradientn(colours=colors,values = scales::rescale(c(-1,-0.6,-0,0.6,1)),limits = c(-1,1))+
          scale_colour_manual(values = c("sig" = "black", "non_sig" = "grey")) +
          scale_size_manual(values = c("sig" = 0.6, "non_sig" = 0.2)) +
          theme_void() +
          theme(title=element_text(size=12),plot.title = element_text(hjust = 0.5),legend.text=element_text(size=6),
                legend.position = "none")+
          theme(axis.line = element_blank(),
                panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                panel.border = element_blank(),
                panel.background = element_blank())
  
  plot2 <- ggseg(.data = data2,atlas = "aseg",view = "coronal",mapping = aes(fill = Effect, colour = highlight,size = highlight),size=1,
                 show.legend = F) +
        scale_fill_gradientn(colours=colors,values = scales::rescale(c(-1,-0.6,-0,0.6,1)),limits = c(-1,1))+
        scale_colour_manual(values = c("sig" = "black", "non_sig" = "grey")) +
        scale_size_manual(values = c("sig" = 0.6, "non_sig" = 0.2)) +
        theme_void() +
        theme(title=element_text(size=12),plot.title = element_text(hjust = 0.5),legend.text=element_text(size=6),
              legend.position = "none")+
        theme(axis.line = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.border = element_blank(),
              panel.background = element_blank())
  
  plot_data_analysis = ggarrange(plot1,plot2,ncol = 1,nrow = 2,heights = c(2,1))
  
  return(plot_data_analysis)
}


# C1_domains <- calculate_cognition_domians_Annebet(C1,C2s)

calculate_cognition_domians_Annebet <- function(C1,C2S){
  
  # Define cognitive measure fields
  em <- c('N_15WT1','N_15WT2','N_15WT3','N_15WT4','N_15WT5','N_15WTui',
           "N_15WTvn","N_15WTvp",'N_VATA1', 'N_VATA2')
  ef <- c('N_FLD', 'N_FLA', 'N_FLT','N_TMTAt','N_TMTBt', 'N_CRspa', 'N_Strt3', 'N_Strt2')
  ap <- c('N_CRspv', 'N_TMTAt', 'N_Strt1')# remove N_LDST_90
  la <- c('N_FL60', 'N_Vatben')
  vf <- c('N_Reykop', 'N_NumLoc', 'N_FraLet') # remove N_Reykop
  
  fields_of_interest <- c('I_ID', 'N_date', unique(c(em, ef, ap, la, vf)))
  
  # Cleaning data in C1 and C2s
  for (c in unique(c(em, ef, ap, la, vf))) {
    C1[[c]][C1[[c]] >= 997] <- NA
    C2s[[c]][C2s[[c]] >= 997] <- NA
  }
  C1$N_VATB3[C1$N_VATB3 == 132] <- NA
  
  # Apply specific logic to fill missing values based on conditions
  fill_logic <- function(df) {
    df$N_VATB2[with(df, N_VATB1 == 6 & is.na(N_VATB2))] <- 6
    df$N_VATA2[with(df, N_VATA1 == 6 & is.na(N_VATA2))] <- 6
    df$N_VATA3[with(df, N_VATA1 == 6 | N_VATA2 == 6)] <- 6
    df$N_VATB3[with(df, N_VATB1 == 6 | N_VATB2 == 6)] <- 6
    return(df)
  }

  C1 <- fill_logic(C1)
  C2s <- fill_logic(C2s)
  
  ############# Memory ############# 
  em_wt15_a <- rowSums(C1[,c('N_15WT1','N_15WT2','N_15WT3','N_15WT4','N_15WT5')], na.rm = TRUE)
  em_wt15_s <- rowSums(C2s[,c('N_15WT1','N_15WT2','N_15WT3','N_15WT4','N_15WT5')], na.rm = TRUE)
  C1$em_wt15 <- (em_wt15_a - mean(em_wt15_s, na.rm = TRUE)) / mean(em_wt15_s, na.rm = TRUE) # Normalizing em1b
  
  em_Delayed_a <- C1$N_15WTui
  em_Delayed_s <- C2s$N_15WTui
  C1$em_Delayed <- (em_Delayed_a - mean(em_Delayed_s, na.rm = TRUE)) / mean(em_Delayed_s, na.rm = TRUE)
  
  em_Recognition_a = 30 - rowSums(C1[,c('N_15WTvn','N_15WTvp')], na.rm = TRUE)
  em_Recognition_s = 30 - rowSums(C2s[,c('N_15WTvn','N_15WTvp')], na.rm = TRUE)
  C1$em_Recognition <- (em_Recognition_a - mean(em_Recognition_s, na.rm = TRUE)) / mean(em_Recognition_s, na.rm = TRUE)
  em_WT = rowMeans(C1[,c('em_wt15', 'em_Delayed',"em_Recognition")], na.rm = TRUE)
  
  em2 <- rowSums(C1[,c('N_VATA1', 'N_VATA2')], na.rm = TRUE)
  em2s <- rowSums(C2s[,c('N_VATA1', 'N_VATA2')], na.rm = TRUE)
  em_VATA <- (em2 - mean(em2s, na.rm = TRUE)) / sd(em2s, na.rm = TRUE)
  C1$EpisodicMemory <- rowMeans(cbind(em_WT, em_VATA), na.rm = TRUE)
  
  
  ############# ExecutiveFunctioning  ############# 
  # N_fluency: N_FLD + N_FLA + N_FLT (=phonetic fluency)
  ef4 <- rowSums(C1[,c('N_FLD', 'N_FLA', 'N_FLT')], na.rm = TRUE)
  ef4s <- rowSums(C2s[,c('N_FLD', 'N_FLA', 'N_FLT')], na.rm = TRUE)
  ef4 <- (ef4 - mean(ef4s, na.rm = TRUE)) / sd(ef4s, na.rm = TRUE)
  
  # TMTB: TMT B/A index (based on N_TMTAt / N_TMTBt)
  ef1 <- (C1$N_TMTBt)/(C1$N_TMTAt)
  ef1s <- (C2s$N_TMTBt)/(C2s$N_TMTAt)
  ef1 <- (ef1 - mean(ef1s, na.rm = TRUE)) / sd(ef1s, na.rm = TRUE)
  
  # Stroop test: time on Stroop card 3 / ([time card 1 + time card 2] / 2)
  ef3 <- C1$N_Strt3/(0.5*rowSums(C1[,c('N_Strt1', 'N_Strt2')], na.rm = TRUE))
  ef3s <- C2s$N_Strt3/(0.5*rowSums(C2s[,c('N_Strt1', 'N_Strt2')], na.rm = TRUE))
  ef3 <- (ef3 - mean(ef3s, na.rm = TRUE)) / sd(ef3s, na.rm = TRUE)
  
  # Digit span, backward: N_CRspa (= working memory)
  ef2 <- C1$N_CRspa
  ef2s <- C2s$N_CRspa
  ef2 <- (ef2 - mean(ef2s, na.rm = TRUE)) / sd(ef2s, na.rm = TRUE)
  # Executive Function:  mean(N_fluency, N_CRspa, TMTB,Stroop test)
  C1$ExecutiveFunctioning <- rowMeans(cbind(ef4,ef1,ef2,ef3), na.rm = TRUE)

  
  ############# Attention and processing speed  ############# 
  ap1 <- (C1$N_CRspv - mean(C2s$N_CRspv, na.rm = TRUE)) / sd(C2s$N_CRspv, na.rm = TRUE)
  ap2 <- -(C1$N_TMTAt - mean(C2s$N_TMTAt, na.rm = TRUE)) / sd(C2s$N_TMTAt, na.rm = TRUE)
  ap3 <- -(C1$N_Strt1 - mean(C2s$N_Strt1, na.rm = TRUE)) / sd(C2s$N_Strt1, na.rm = TRUE)
  ap4 <- -(C1$N_Strt2 - mean(C2s$N_Strt2, na.rm = TRUE)) / sd(C2s$N_Strt2, na.rm = TRUE)
  ap5 <- (C1$N_LDST_90 - mean(C2s$N_LDST_90, na.rm = TRUE)) / sd(C2s$N_LDST_90, na.rm = TRUE)
  C1$Attention <- rowMeans(cbind(ap1,ap2,ap3,ap4), na.rm = TRUE)

  #############  Language: mean(N_FL60, N_Vatben) ############# 
  la1 <- (C1$N_FL60 - mean(C2s$N_FL60, na.rm = TRUE)) / sd(C2s$N_FL60, na.rm = TRUE)
  la2 <- (C1$N_Vatben - mean(C2s$N_Vatben, na.rm = TRUE)) / sd(C2s$N_Vatben, na.rm = TRUE)
  C1$Language <- rowMeans(cbind(la1, la2), na.rm = TRUE)

  #############  Visualspatial functioning: mean(N_Reykop, N_NumLoc, N_FraLet) ############# 
  vf1 <- (C1$N_Reykop - mean(C2s$N_Reykop, na.rm = TRUE)) / sd(C2s$N_Reykop, na.rm = TRUE)
  vf2 <- (C1$N_NumLoc - mean(C2s$N_NumLoc, na.rm = TRUE)) / sd(C2s$N_NumLoc, na.rm = TRUE)
  vf3 <- (C1$N_FraLet - mean(C2s$N_FraLet, na.rm = TRUE)) / sd(C2s$N_FraLet, na.rm = TRUE)
  C1$VisuospatialFunctioning <- rowMeans(cbind(vf2, vf3), na.rm = TRUE)

  return(C1)
}

calculate_cognition_domians_Annebet <- function(C1,C2S){
  
  # Define cognitive measure fields
  em <- c('N_15WT1','N_15WT2','N_15WT3','N_15WT4','N_15WT5','N_15WTui',
           "N_15WTvn","N_15WTvp",'N_VATA1', 'N_VATA2')
  ef <- c('N_FLD', 'N_FLA', 'N_FLT','N_TMTAt','N_TMTBt', 'N_CRspa', 'N_Strt3', 'N_Strt2')
  ap <- c('N_CRspv', 'N_TMTAt', 'N_Strt1')# remove N_LDST_90
  la <- c('N_FL60', 'N_Vatben')
  vf <- c('N_Reykop', 'N_NumLoc', 'N_FraLet') # remove N_Reykop
  
  fields_of_interest <- c('I_ID', 'N_date', unique(c(em, ef, ap, la, vf)))
  
  # Cleaning data in C1 and C2s
  for (c in unique(c(em, ef, ap, la, vf))) {
    C1[[c]][C1[[c]] >= 997] <- NA
    C2s[[c]][C2s[[c]] >= 997] <- NA
  }
  C1$N_VATB3[C1$N_VATB3 == 132] <- NA
  
  # Apply specific logic to fill missing values based on conditions
  fill_logic <- function(df) {
    df$N_VATB2[with(df, N_VATB1 == 6 & is.na(N_VATB2))] <- 6
    df$N_VATA2[with(df, N_VATA1 == 6 & is.na(N_VATA2))] <- 6
    df$N_VATA3[with(df, N_VATA1 == 6 | N_VATA2 == 6)] <- 6
    df$N_VATB3[with(df, N_VATB1 == 6 | N_VATB2 == 6)] <- 6
    return(df)
  }

  C1 <- fill_logic(C1)
  C2s <- fill_logic(C2s)
  
  ############# Memory ############# 
  em_wt15_a <- rowSums(C1[,c('N_15WT1','N_15WT2','N_15WT3','N_15WT4','N_15WT5')], na.rm = TRUE)
  em_wt15_s <- rowSums(C2s[,c('N_15WT1','N_15WT2','N_15WT3','N_15WT4','N_15WT5')], na.rm = TRUE)
  C1$em_wt15 <- (em_wt15_a - mean(em_wt15_s, na.rm = TRUE)) / mean(em_wt15_s, na.rm = TRUE) # Normalizing em1b
  
  em_Delayed_a <- C1$N_15WTui
  em_Delayed_s <- C2s$N_15WTui
  C1$em_Delayed <- (em_Delayed_a - mean(em_Delayed_s, na.rm = TRUE)) / mean(em_Delayed_s, na.rm = TRUE)
  
  em_Recognition_a = 30 - rowSums(C1[,c('N_15WTvn','N_15WTvp')], na.rm = TRUE)
  em_Recognition_s = 30 - rowSums(C2s[,c('N_15WTvn','N_15WTvp')], na.rm = TRUE)
  C1$em_Recognition <- (em_Recognition_a - mean(em_Recognition_s, na.rm = TRUE)) / mean(em_Recognition_s, na.rm = TRUE)
  em_WT = rowMeans(C1[,c('em_wt15', 'em_Delayed',"em_Recognition")], na.rm = TRUE)
  
  em2 <- rowSums(C1[,c('N_VATA1', 'N_VATA2')], na.rm = TRUE)
  em2s <- rowSums(C2s[,c('N_VATA1', 'N_VATA2')], na.rm = TRUE)
  em_VATA <- (em2 - mean(em2s, na.rm = TRUE)) / sd(em2s, na.rm = TRUE)
  C1$EpisodicMemory <- rowMeans(cbind(em_WT, em_VATA), na.rm = TRUE)
  
  
  ############# ExecutiveFunctioning  ############# 
  # N_fluency: N_FLD + N_FLA + N_FLT (=phonetic fluency)
  ef4 <- rowSums(C1[,c('N_FLD', 'N_FLA', 'N_FLT')], na.rm = TRUE)
  ef4s <- rowSums(C2s[,c('N_FLD', 'N_FLA', 'N_FLT')], na.rm = TRUE)
  ef4 <- (ef4 - mean(ef4s, na.rm = TRUE)) / sd(ef4s, na.rm = TRUE)
  
  # TMTB: TMT B/A index (based on N_TMTAt / N_TMTBt)
  ef1 <- (C1$N_TMTBt)/(C1$N_TMTAt)
  ef1s <- (C2s$N_TMTBt)/(C2s$N_TMTAt)
  ef1 <- (ef1 - mean(ef1s, na.rm = TRUE)) / sd(ef1s, na.rm = TRUE)
  
  # Stroop test: time on Stroop card 3 / ([time card 1 + time card 2] / 2)
  ef3 <- C1$N_Strt3/(0.5*rowSums(C1[,c('N_Strt1', 'N_Strt2')], na.rm = TRUE))
  ef3s <- C2s$N_Strt3/(0.5*rowSums(C2s[,c('N_Strt1', 'N_Strt2')], na.rm = TRUE))
  ef3 <- (ef3 - mean(ef3s, na.rm = TRUE)) / sd(ef3s, na.rm = TRUE)
  
  # Digit span, backward: N_CRspa (= working memory)
  ef2 <- C1$N_CRspa
  ef2s <- C2s$N_CRspa
  ef2 <- (ef2 - mean(ef2s, na.rm = TRUE)) / sd(ef2s, na.rm = TRUE)
  # Executive Function:  mean(N_fluency, N_CRspa, TMTB,Stroop test)
  C1$ExecutiveFunctioning <- rowMeans(cbind(ef4,ef1,ef2,ef3), na.rm = TRUE)

  
  ############# Attention and processing speed  ############# 
  ap1 <- (C1$N_CRspv - mean(C2s$N_CRspv, na.rm = TRUE)) / sd(C2s$N_CRspv, na.rm = TRUE)
  ap2 <- -(C1$N_TMTAt - mean(C2s$N_TMTAt, na.rm = TRUE)) / sd(C2s$N_TMTAt, na.rm = TRUE)
  ap3 <- -(C1$N_Strt1 - mean(C2s$N_Strt1, na.rm = TRUE)) / sd(C2s$N_Strt1, na.rm = TRUE)
  ap4 <- -(C1$N_Strt2 - mean(C2s$N_Strt2, na.rm = TRUE)) / sd(C2s$N_Strt2, na.rm = TRUE)
  ap5 <- (C1$N_LDST_90 - mean(C2s$N_LDST_90, na.rm = TRUE)) / sd(C2s$N_LDST_90, na.rm = TRUE)
  C1$Attention <- rowMeans(cbind(ap1,ap2,ap3,ap4), na.rm = TRUE)

  #############  Language: mean(N_FL60, N_Vatben) ############# 
  la1 <- (C1$N_FL60 - mean(C2s$N_FL60, na.rm = TRUE)) / sd(C2s$N_FL60, na.rm = TRUE)
  la2 <- (C1$N_Vatben - mean(C2s$N_Vatben, na.rm = TRUE)) / sd(C2s$N_Vatben, na.rm = TRUE)
  C1$Language <- rowMeans(cbind(la1, la2), na.rm = TRUE)

  #############  Visualspatial functioning: mean(N_Reykop, N_NumLoc, N_FraLet) ############# 
  vf1 <- (C1$N_Reykop - mean(C2s$N_Reykop, na.rm = TRUE)) / sd(C2s$N_Reykop, na.rm = TRUE)
  vf2 <- (C1$N_NumLoc - mean(C2s$N_NumLoc, na.rm = TRUE)) / sd(C2s$N_NumLoc, na.rm = TRUE)
  vf3 <- (C1$N_FraLet - mean(C2s$N_FraLet, na.rm = TRUE)) / sd(C2s$N_FraLet, na.rm = TRUE)
  C1$VisuospatialFunctioning <- rowMeans(cbind(vf2, vf3), na.rm = TRUE)

  return(C1)
}


missing_report <- function(data,column_predictor){
  
  Y <- data[,column_predictor]
  missing <- round(colSums(is.na(Y) / nrow(Y)) * 100, 1)
  return(missing)
}



mutate_group_outliers_na <- function(.data, group_col, var_cols,
                                     sd_thresh = 3, na.rm = TRUE) {
  .data %>%
    group_by({{ group_col }}) %>%
    mutate(
      across(
        all_of(var_cols),
        ~ replace(
            .,
            abs(. - mean(., na.rm = na.rm)) > sd_thresh * sd(., na.rm = na.rm),
            NA_real_
          ),
        .names = "{.col}"
      )
    ) %>%
    ungroup()
}



mutate_group_outliers_iqr_na <- function(.data, group_col, var_cols,
                                         iqr_thresh = 1.5, na.rm = TRUE) {
  .data %>%
    group_by({{ group_col }}) %>%
    mutate(
      across(
        all_of(var_cols),
        ~ {
          # compute 25th and 75th percentiles
          qs  <- quantile(., probs = c(0.25, 0.75), na.rm = na.rm)
          Q1  <- qs[1]; Q3 <- qs[2]
          iqr <- Q3 - Q1
  
          # define lower/upper fences
          lower <- Q1 - iqr_thresh * iqr
          upper <- Q3 + iqr_thresh * iqr
  
          # replace outside fence with NA
          replace(., . < lower | . > upper, NA_real_)
        },
        .names = "{.col}"
      )
    ) %>%
    ungroup()
}
```


# general data clean 
```{r,include=FALSE}

# MRI phenotypes
names_CT <- c(names(Freesurfer_output_718)[165:199],names(Freesurfer_output_718)[202:236])
names_Vol <- c(names(Freesurfer_output_718)[240:273],names(Freesurfer_output_718)[276:309])

names_Sub_1<- c("Left-Thalamus","Left-Caudate","Left-Putamen","Left-Pallidum","Left-Hippocampus","Left-Amygdala",
              "Left-Accumbens-area","Left-choroid-plexus",
              "Right-Thalamus","Right-Caudate","Right-Putamen","Right-Pallidum","Right-Hippocampus","Right-Amygdala",
              "Right-Accumbens-area","Right-choroid-plexus")
names_Sub_2<- c("lh_Thalamus_volume","lh_Caudate_volume","lh_Putamen_volume","lh_Pallidum_volume","lh_Hippocampus_volume","lh_Amygdala_volume",
              "lh_Accumbens_area_volume","lh_choroid_plexus_volume",
              "rh_Thalamus_volume","rh_Caudate_volume","rh_Putamen_volume","rh_Pallidum_volume","rh_Hippocampus_volume","rh_Amygdala_volume",
              "rh_Accumbens_area_volume","rh_choroid_plexus_volume")
names_eTIV <- c("EstimatedTotalIntraCranialVol")
MRI_IDPs = c(names_CT,names_Vol,names_Sub_1,names_eTIV)
MRI_IDPs_after = c(names_CT,names_Vol,names_Sub_2,"eTIV")


data_718_MRI_v00 <- data_718_MRI %>% 
  filter(!(MRI_ID == "NULL" | is.na(M_dat)))

data_718_MRI_v0 <- Freesurfer_output_718 %>% 
  select(MRI_ID,I_ID,ManufacturersModelName,MagneticFieldStrength,MRI_IDPs) %>%
  mutate(MRI_ID = str_split(MRI_ID,"sub-",simplify = TRUE)[,2]) %>%
  inner_join(data_718_MRI_v00[,c("MRI_ID","M_dat","M_age","M_scan","M_quality","M_Fazekas")],by = "MRI_ID") %>%
  mutate(M_dat = ymd(M_dat)) %>%
  mutate(I_ID = as.numeric(I_ID)) %>%
  filter(!if_any(all_of(c(MRI_IDPs)), is.na)) %>%
  
  group_by(ManufacturersModelName) %>%
  filter(n() >= 20) %>%
  ungroup() %>%
  
  filter(!is.na(I_ID) & !is.na(M_dat)) %>%
  
  mutate(MagneticFieldStrength = case_when(
    MagneticFieldStrength == 0.95007 ~ 1,
    MagneticFieldStrength == 1.494 ~ 1.5,
    MagneticFieldStrength == 1.5 ~ 1.5,
    MagneticFieldStrength == 3 ~ 3
  )) %>%
  filter(!(ManufacturersModelName == "SIGNA HDx" & MagneticFieldStrength == 1.5))


data_718_Baseline_v0 <- data_718_Baseline %>% mutate(I_dat_1 = ymd(I_dat_1)) %>%
  filter(!I_sex == "NULL" & !I_edu_VE == "NULL") %>%
  mutate(APOE_status = case_when(
    APOE %in% c("E2E4","E3E4","E4E4")~"e4-carrier",
    APOE %in% c("E2E2","E2E3","E3E3")~"e4-non-carrier",
    APOE == c("NULL")~"Missing",
  )) %>%
  mutate(Edu = case_when(
    I_edu_VE == 1 ~ 5,
    I_edu_VE == 2 ~ 6,
    I_edu_VE == 3 ~ 8,
    I_edu_VE == 4 ~ 9,
    I_edu_VE == 5 ~ 10,
    I_edu_VE == 6 ~ 13,
    I_edu_VE == 7 ~ 17
  )) %>%
  mutate(Sex = if_else(I_sex %in% c("M","m"),"m","f")) %>%
  mutate(Sex = factor(Sex,c("m","f"))) %>%
  select(I_ID,Sex,Edu,APOE_status)


data_718_Diagnosis_v0 <- data_718_Diagnosis %>% 
  mutate(D_dat = ymd(D_dat)) %>%
  #filter(diagnoseOms %in% c("Subjectieve klachten","MCI","Possible AD","Probable AD","FTD","DLB")) %>%
  mutate(D_diagnosis = case_when(
  diagnoseOms == "Subjectieve klachten"~"SCD",
  diagnoseOms == "MCI"~"MCI",
  diagnoseOms == "Probable AD"~"AD",
  diagnoseOms == "FTD"~"FTD",
  diagnoseOms == "DLB"~"DLB",
  TRUE ~ diagnoseOms,
))


data_718_CSF_v0 <- data_718_CSF %>% 
  mutate(L_DATEAFN = ymd(L_DATEAFN)) %>%
  filter(!L_AB42_corr == "NULL" | !L_AB42_Elecsys == "NULL") %>%
  mutate(ATN_AB = case_when(is.na(L_AB42_corr) & is.na(L_AB42_Elecsys) ~ NA,
                            L_AB42_Elecsys < 1000 & !is.na(L_AB42_Elecsys) ~ "A+",
                            L_AB42_Elecsys >= 1000 & !is.na(L_AB42_Elecsys) ~ "A-",
                            L_AB42_corr < 813 & is.na(L_AB42_Elecsys) ~ "A+",
                            L_AB42_corr >= 813 & is.na(L_AB42_Elecsys) ~ "A-")) %>%
  mutate(across(c("L_AB42_Elecsys","L_AB42_corr","L_TAU_Elecsys","L_TAU","L_PTAU_Elecsys","L_PTAU"), as.numeric)) %>%
  mutate(L_AB42_norm = ifelse(!is.na(L_AB42_Elecsys), L_AB42_Elecsys, (-365 + 1.87 * L_AB42_corr)), 
         L_TAU_norm = ifelse(!is.na(L_TAU_Elecsys), L_TAU_Elecsys, (-365 + 1.87 * L_TAU)), 
         L_PTAU_norm = ifelse(!is.na(L_PTAU_Elecsys), L_PTAU_Elecsys, -3.9 + 0.44 * L_PTAU)) %>%
  filter(!if_any(all_of(c("L_AB42_norm","L_TAU_norm","L_PTAU_norm")), is.na))
  

data_718_PET_v0 <- data_718_PET %>% mutate(PT_date = ymd(PT_date)) %>% 
  filter(!is.na(PT_result_amyloid))


data_718_NPO_v0 <- data_718_NPO %>% 
  mutate(N_date = ymd(N_date)) %>% 
  mutate(across(where(is.character), ~na_if(., "NULL"))) %>%
  mutate(across(starts_with("N_") & !any_of(c("N_date", "N_age")), as.numeric)) %>%
  mutate(across(
    starts_with("N_") & !any_of(c("N_date", "N_age")), 
    ~ifelse(. >= 997, NA, .)))

data_718_MMSE_v0 <- data_718_MMSE %>% mutate(V_dat = ymd(V_dat)) %>% mutate(V_MMSE = as.numeric(V_MMSE)) %>% filter(!is.na(V_MMSE))
```



# Main Data Selection
```{r}
############################## harmonize CSF and PET the "bind_rows" later ############################## 
data_718_CSF_v1 <- data_718_CSF_v0 %>% select(I_ID,L_DATEAFN,CSF_A,L_age) %>%
  rename(amyloid_DATEAFN = L_DATEAFN,amyloid_status = CSF_A) %>%
  mutate(amyloid_source = "CSF")

data_718_PET_v1 <- data_718_PET_v0 %>% select(I_ID,PT_date,PT_result_amyloid) %>%
  rename(amyloid_DATEAFN = PT_date,amyloid_status = PT_result_amyloid) %>%
  mutate(amyloid_status = if_else(amyloid_status %in% c("positief","Positief"),"A+","A-")) %>%
  mutate(amyloid_source = "PET")

############################## data selection based on CSF ############################## 
data_718_CSF_unique <- data_718_CSF_v1 %>% 
  
  inner_join(data_718_Baseline_v0,by = "I_ID",relationship = "many-to-many") %>% 
  
  inner_join(data_718_MRI_v0,by = "I_ID",relationship = "many-to-many") %>%
  mutate(diff_L_M_raw = difftime(amyloid_DATEAFN,M_dat,units ="days"),diff_L_M = abs(diff_L_M_raw),relationship = "many-to-many") %>%
  
  inner_join(data_718_NPO_v0,by = "I_ID",relationship = "many-to-many") %>%
  mutate(diff_L_N_raw = difftime(amyloid_DATEAFN,N_date,units ="days"),diff_L_N = abs(diff_L_N_raw),relationship = "many-to-many") %>%
  
  inner_join(data_718_Diagnosis_v0,by = "I_ID",relationship = "many-to-many") %>%
  mutate(diff_L_D = difftime(D_dat,amyloid_DATEAFN,units ="days"),relationship = "many-to-many") %>% 
  
  inner_join(data_718_MMSE,by = "I_ID",relationship = "many-to-many") %>%
  mutate(diff_L_V_raw = difftime(amyloid_DATEAFN,V_dat,units ="days"),diff_L_V = abs(diff_L_V_raw),relationship = "many-to-many") %>%
  
  filter((abs(diff_L_M) <= 365 & abs(diff_L_N) <= 365) & abs(diff_L_D) <= 365 & abs(diff_L_V) <= 365) %>%
  
  group_by(I_ID) %>%
  filter(diff_L_D == max(diff_L_D)) %>%
  arrange(diff_L_M,diff_L_N,diff_L_V) %>%
  slice(1) %>%
  ungroup()


##############################  data selection based on PET ############################## 
data_718_PET_unique <- data_718_PET_v1 %>% 
  
  inner_join(data_718_Baseline_v0,by = "I_ID",relationship = "many-to-many") %>% 
  
  inner_join(data_718_MRI_v0,by = "I_ID",relationship = "many-to-many") %>%
  mutate(diff_L_M_raw = difftime(amyloid_DATEAFN,M_dat,units ="days"),diff_L_M = abs(diff_L_M_raw),relationship = "many-to-many") %>%
  
  inner_join(data_718_NPO_v0,by = "I_ID",relationship = "many-to-many") %>%
  mutate(diff_L_N_raw = difftime(amyloid_DATEAFN,N_date,units ="days"),diff_L_N = abs(diff_L_N_raw),relationship = "many-to-many") %>%
  
  inner_join(data_718_Diagnosis_v0,by = "I_ID",relationship = "many-to-many") %>%
  mutate(diff_L_D = difftime(D_dat,amyloid_DATEAFN,units ="days"),relationship = "many-to-many") %>% 
  
  inner_join(data_718_MMSE,by = "I_ID",relationship = "many-to-many") %>%
  mutate(diff_L_V_raw = difftime(amyloid_DATEAFN,V_dat,units ="days"),diff_L_V = abs(diff_L_V_raw),relationship = "many-to-many") %>%
  
  filter((abs(diff_L_M) <= 365 & abs(diff_L_N) <= 365) & abs(diff_L_D) <= 365 & abs(diff_L_V) <= 365) %>%
  
  group_by(I_ID) %>%
  filter(diff_L_D == max(diff_L_D)) %>%
  arrange(diff_L_M,diff_L_N,diff_L_V) %>%
  slice(1) %>%
  ungroup()


############################## bind_rows two DF based on rules: PET first ############################## 
data_ADC_analysis <- bind_rows(data_718_PET_unique,data_718_CSF_unique) %>%
  group_by(I_ID) %>%
  filter(!(amyloid_source == "CSF" & any(amyloid_source == "PET"))) %>%
  ungroup() %>%
  arrange(I_ID,amyloid_DATEAFN) %>%
  filter(D_diagnosis %in% c("SCD","AD","DLB","FTD")) %>%
  filter(!(D_diagnosis == "SCD"& amyloid_status == "A+")) %>%
  filter(!(D_diagnosis == "AD"& amyloid_status == "A-")) %>%
  mutate(amyloid_age = if_else(
    amyloid_source == "PET", M_age + round(as.numeric(difftime(amyloid_DATEAFN,M_dat,units ="days"))/365,2),L_age))


# dim(data_ADC_analysis)
# table(data_ADC_analysis$D_diagnosis,data_ADC_analysis$amyloid_status)
# data_ADC_analysis_dip <- data_ADC_analysis %>% filter(I_ID %in% data_ADC_analysis$I_ID[duplicated(data_ADC_analysis$I_ID)])

data_ADC_analysis <- data_ADC_analysis %>%
  mutate_group_outliers_na(
    group_col = D_diagnosis,
    var_cols  = c("N_Strt1","N_Strt2","N_Strt3","N_TMTBt","N_TMTAt"),
    sd_thresh = 3
  )


#amyloid_compare <- data_718_PET_unique %>%
#  filter(D_diagnosis %in% c("SCD","AD","DLB","FTD")) %>%
#  select(I_ID, PET_status = amyloid_status) %>%
#  inner_join(
#    data_718_CSF_unique %>%
#      select(I_ID, CSF_status = amyloid_status,D_diagnosis) %>% filter(D_diagnosis %in% c("SCD","AD","DLB","FTD")),
#    by = "I_ID"
#  )

#table(amyloid_compare$CSF_status, amyloid_compare$PET_status)

#X <- amyloid_compare %>% filter(!CSF_status == PET_status)
#table(X$D_diagnosis)
```


# MRI data preprocessing: correct for scanner effects
```{r}
# MRI phenotypes
names_CT <- c(names(Freesurfer_output_718)[165:199],names(Freesurfer_output_718)[202:236])
names_Vol <- c(names(Freesurfer_output_718)[240:273],names(Freesurfer_output_718)[276:309])

names_Sub_1<- c("Left-Thalamus","Left-Caudate","Left-Putamen","Left-Pallidum","Left-Hippocampus","Left-Amygdala",
              "Left-Accumbens-area","Left-choroid-plexus",
              "Right-Thalamus","Right-Caudate","Right-Putamen","Right-Pallidum","Right-Hippocampus","Right-Amygdala",
              "Right-Accumbens-area","Right-choroid-plexus")
names_Sub_2<- c("lh_Thalamus_volume","lh_Caudate_volume","lh_Putamen_volume","lh_Pallidum_volume","lh_Hippocampus_volume","lh_Amygdala_volume",
              "lh_Accumbens_area_volume","lh_choroid_plexus_volume",
              "rh_Thalamus_volume","rh_Caudate_volume","rh_Putamen_volume","rh_Pallidum_volume","rh_Hippocampus_volume","rh_Amygdala_volume",
              "rh_Accumbens_area_volume","rh_choroid_plexus_volume")
names_eTIV <- c("EstimatedTotalIntraCranialVol")
MRI_IDPs = c(names_CT,names_Vol,names_Sub_1,names_eTIV)
MRI_IDPs_after = c(names_CT,names_Vol,names_Sub_2,"eTIV")

################################## correction scanner * sites
data_ADC_analysis_MRI <- data_ADC_analysis %>%
  select(MRI_ID, M_age, Sex, Edu, D_diagnosis,amyloid_status,ManufacturersModelName, MagneticFieldStrength,EstimatedTotalIntraCranialVol,all_of(MRI_IDPs)) %>% 
  rename(Scanner = ManufacturersModelName, Field_Strength = MagneticFieldStrength) %>%
  mutate(eTIV = EstimatedTotalIntraCranialVol)

clindata_path_data = data_ADC_analysis_MRI[,c("MRI_ID","M_age","Sex","Edu","D_diagnosis")]
mridata_path_data = data_ADC_analysis_MRI

writexl::write_xlsx(clindata_path_data,"/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/data/MRI_scanner/clindata_path_data.xlsx")
writexl::write_xlsx(mridata_path_data,"/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/data/MRI_scanner/mridata_path_data.xlsx")

# python /Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/scripts/combat_ADC_mega.py

data_ADC_analysis_MRI_filtered = read_xlsx("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/data/MRI_scanner/final_23_June2025.xlsx")
data_ADC_analysis_v1 <- data_ADC_analysis %>% 
  select(-all_of(c(names_CT,names_Vol,names_Sub_1,names_eTIV))) %>%
  left_join(data_ADC_analysis_MRI_filtered[,c("MRI_ID",c(names_CT,names_Vol,names_Sub_1,"eTIV"))],by = c("MRI_ID")) %>%
  rename_with(.cols = all_of(names_Sub_1),
              .fn = ~ str_replace_all(., "-", "_")) %>%
  rename_with(.cols = starts_with("Left_"),
              .fn = ~ str_replace_all(., "Left_", "lh_") %>% paste0("_volume")) %>%
  rename_with(.cols = starts_with("Right_"),
              .fn = ~ str_replace_all(., "Right_", "rh_") %>% paste0("_volume")) %>%
  
  mutate(across(starts_with("lh_"), 
                .fns = list(mean = ~ ( . + get(sub("lh_", "rh_", cur_column())) ) / 2), 
                .names = "{sub('lh_', '', .col)}_mean")) %>% 
  mutate(across(starts_with("Left."), 
                .fns = list(mean = ~ ( . + get(sub("Left.", "Right.", cur_column())) ) / 2), 
                .names = "{sub('Left.', '', .col)}_mean")) %>%
  
  ### scaling
  mutate(across(all_of(c(names_CT,names_Vol,names_Sub_2,"eTIV")), scale)) %>%
  mutate(across(ends_with("_mean"), scale))


save(data_ADC_analysis_v1,file = "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/scripts/Pati_Combat/ADC_harmonized.RData")
```


############ combat visualization plot
```{r}
data_ADC_analysis_MRI_filtered <- data_ADC_analysis_MRI_filtered %>% 
  inner_join(data_ADC_analysis_MRI[,c("MRI_ID","Scanner","Field_Strength","D_diagnosis","M_age")],"MRI_ID")

data_ADC_analysis_MRI$Left_Hippocampus = data_ADC_analysis_MRI$`Left-Hippocampus`
data_ADC_analysis_MRI$Field_Strength = as.character(data_ADC_analysis_MRI$Field_Strength)
data_ADC_analysis_MRI$scaner_type = str_c(data_ADC_analysis_MRI$Scanner,":",data_ADC_analysis_MRI$Field_Strength,"T")
data_ADC_analysis_MRI$scaner_type = factor(data_ADC_analysis_MRI$scaner_type,c("Signa HDxt:3T","Avanto:1.5T","DISCOVERY MR750:3T","Ingenuity:3T","MAGNETOM EXPERT:1T","MAGNETOM Vida:3T","SIGNA HDx:3T","Signa HDxt:1.5T","Sonata:1.5T","Titan3T:3T"))

data_ADC_analysis_MRI_filtered$Left_Hippocampus = data_ADC_analysis_MRI_filtered$`Left-Hippocampus`
data_ADC_analysis_MRI_filtered$Field_Strength = as.character(data_ADC_analysis_MRI_filtered$Field_Strength)
data_ADC_analysis_MRI_filtered$scaner_type = str_c(data_ADC_analysis_MRI_filtered$Scanner,":",data_ADC_analysis_MRI_filtered$Field_Strength,"T")
data_ADC_analysis_MRI_filtered$scaner_type = factor(data_ADC_analysis_MRI_filtered$scaner_type,c("Signa HDxt:3T","Avanto:1.5T","DISCOVERY MR750:3T","Ingenuity:3T","MAGNETOM EXPERT:1T","MAGNETOM Vida:3T","SIGNA HDx:3T","Signa HDxt:1.5T","Sonata:1.5T","Titan3T:3T"))
## ============================================
## Scanner-type effects on Left Hippocampus
## Raw vs ComBat-corrected
## ============================================

library(dplyr)
library(ggplot2)
library(broom)
library(scales)   # for pvalue()
library(ggpubr)   # for ggarrange

## ------------------------------------------------
## 0. Helper: function to get scanner forest data
## ------------------------------------------------

scanner_forest_df <- function(dat) {
  
  fit <- lm(Left_Hippocampus ~ scaner_type + M_age + Sex + Edu + D_diagnosis + eTIV, data = dat)

  tidy(fit, conf.int = TRUE) %>%
    filter(grepl("^scaner_type", term)) %>%
    mutate(
      Scanner  = sub("^scaner_type", "", term),
      Beta     = estimate,
      CI_lower = conf.low,
      CI_upper = conf.high,
      p_lab    = pvalue(p.value, accuracy = 1e-2),
      CI_mid   = Beta
    ) %>%
    mutate(Scanner = factor(c("Avanto:1.5T","DISCOVERY MR750:3T","Ingenuity:3T","MAGNETOM EXPERT:1T","MAGNETOM Vida:3T","SIGNA HDx:3T","Signa HDxt:1.5T","Sonata:1.5T","Titan3T:3T"), levels = Scanner))
    #arrange(Beta) %>%
    
}




## ------------------------------------------------
## 1. Prepare forest data (raw & ComBat) and axis
## ------------------------------------------------

df_raw    <- scanner_forest_df(data_ADC_analysis_MRI)
  
df_combat <- scanner_forest_df(data_ADC_analysis_MRI_filtered)

# Get overall CI range across both datasets
ylim_all <- range(
  df_raw$CI_lower, df_raw$CI_upper,
  df_combat$CI_lower, df_combat$CI_upper,
  na.rm = TRUE
)

# Make it symmetric around 0 and a bit wider
max_abs  <- max(abs(ylim_all))
ylim_sym <- c(-max_abs, max_abs) * 1.1   # 10% extra margin


## =========================
## 2. RAW (uncorrected)
## =========================

## 2a. Boxplot – raw

box_raw <- ggplot(
  data_ADC_analysis_MRI,
  aes(x = scaner_type,
      y = Left_Hippocampus,
      group = scaner_type,
      fill  = scaner_type)
) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.5) +
  labs(
    x = NULL,
    y = "Left hippocampal volume",  # add unit if you have it, e.g. (mm^3)
    fill = "Scanner type",
    title = "Raw (uncorrected)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.text.x        = element_text(angle = 45, hjust = 1),
    legend.position    = "NULL",
    legend.title       = element_text(size = 10),
    legend.text        = element_text(size = 9),
    plot.title         = element_text(face = "bold", hjust = 0),
    plot.margin        = margin(5, 5, 5, 5)
  )

## 2b. Forest – raw

forest_raw <- ggplot(
  df_raw,
  aes(x = Scanner, y = Beta, ymin = CI_lower, ymax = CI_upper)
) +
  geom_hline(yintercept = 0, linetype = "dashed",
             color = "grey60", linewidth = 0.4) +
  geom_pointrange(linewidth = 0.5, color = "black") +
  geom_text(
    aes(y = CI_mid, label = p_lab, group = 1),
    position = position_nudge(x = -0.18),
    vjust = 1.4, size = 3.2, lineheight = 0.9, color = "black"
  ) +
  coord_flip(clip = "off") +
  scale_y_continuous(limits = ylim_sym) +
  labs(
    x = NULL,
    y = expression(beta~"(95% CI)"),
    title = "Scanner-type effects (raw)"
  ) +
  theme_classic(base_size = 12) +
  theme(
    panel.grid.major.x = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.minor   = element_blank(),
    axis.text.y        = element_text(size = 10, face = "bold"),
    legend.position    = "none",
    plot.title         = element_text(face = "bold", hjust = 0),
    plot.margin        = margin(5, 15, 5, 5)
  )

## 2c. Combine raw figure

fig_raw <- ggarrange(
  box_raw,
  forest_raw,
  ncol    = 1,
  heights = c(1, 1.1)
)

## 2d. Save raw figure (change format if needed: .pdf, .tiff)

ggsave(
  filename = "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/plots/combat_QC/Figure_scanner_raw.tiff",
  plot     = fig_raw,
  width    = 6, height = 8, dpi = 300
)


## =========================
## 3. ComBat-corrected
## =========================

## 3a. Boxplot – ComBat

box_combat <- ggplot(
  data_ADC_analysis_MRI_filtered,
  aes(x = scaner_type,
      y = Left_Hippocampus,
      group = scaner_type,
      fill  = scaner_type)
) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.5) +
  labs(
    x = NULL,
    y = "Left hippocampal volume",
    fill = "Scanner type",
    title = "ComBat-corrected"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.text.x        = element_text(angle = 45, hjust = 1),
    legend.position    = "NULL",
    legend.title       = element_text(size = 10),
    legend.text        = element_text(size = 9),
    plot.title         = element_text(face = "bold", hjust = 0),
    plot.margin        = margin(5, 5, 5, 5)
  )

## 3b. Forest – ComBat

forest_combat <- ggplot(
  df_combat,
  aes(x = Scanner, y = Beta, ymin = CI_lower, ymax = CI_upper)
) +
  geom_hline(yintercept = 0, linetype = "dashed",
             color = "grey60", linewidth = 0.4) +
  geom_pointrange(linewidth = 0.5, color = "black") +
  geom_text(
    aes(y = CI_mid, label = p_lab, group = 1),
    position = position_nudge(x = -0.18),
    vjust = 1.4, size = 3.2, lineheight = 0.9, color = "black"
  ) +
  coord_flip(clip = "off") +
  scale_y_continuous(limits = ylim_sym) +
  labs(
    x = NULL,
    y = expression(beta~"(95% CI)"),
    title = "Scanner-type effects (ComBat-corrected)"
  ) +
  theme_classic(base_size = 12) +
  theme(
    panel.grid.major.x = element_line(color = "grey90", linewidth = 0.3),
    panel.grid.minor   = element_blank(),
    axis.text.y        = element_text(size = 10, face = "bold"),
    legend.position    = "none",
    plot.title         = element_text(face = "bold", hjust = 0),
    plot.margin        = margin(5, 15, 5, 5)
  )

## 3c. Combine ComBat figure

fig_combat <- ggarrange(
  box_combat,
  forest_combat,
  ncol    = 1,
  heights = c(1, 1.1),
  legend = NULL
)

## 3d. Save ComBat figure

ggsave(
  filename = "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/plots/combat_QC/Figure_scanner_ComBat.tiff",
  plot     = fig_combat,
  width    = 6, height = 8, dpi = 300
)

```


# Annebet's pipeline + without imputation
```{r}
load("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/scripts/Pati_Combat/ADC_harmonized.RData")
data_tmp <- data_ADC_analysis_v1 %>% 
  
  ## memory
  rowwise() %>%
  mutate(
    n_me_WT_sum = log(sum(c_across(N_15WT1:N_15WT5)) + 1),
    n_me_delayed = log(N_15WTui + 1),
    n_me_recognition = log(30 - N_15WTvn - N_15WTvp + 1),
    n_me_vata = log(sum(c_across(N_VATA1:N_VATA2)) + 1),
  ) %>%
  ungroup() %>%

  ## executive functioning
  rowwise() %>%
  mutate(
    n_ef_FL_sum = log(sum(c_across(N_FLD:N_FLT)) + 1),
    n_ef_TMT_index = -1 * log(N_TMTBt/N_TMTAt + 1),
    n_ef_Strt_index = -1 * log(N_Strt3/sum(c_across(c("N_Strt1","N_Strt2"))) + 1)
  ) %>%
  ungroup() %>%

  ## Attention and proicessing speed
  rowwise() %>%
  mutate(
    n_ap_CRspv = log(N_CRspv + 1),
    n_ap_TMTat = -1 * log(N_TMTAt + 1),
    n_ap_Strt1 = -1 * log(N_Strt1 + 1),
    n_ap_Strt2 = -1 * log(N_Strt2 + 1),
    n_ap_LDST_90 = N_LDST_90,
  ) %>%
  ungroup() %>%

  ## Language
  rowwise() %>%
  mutate(
    n_la_FL60 = log(N_FL60 + 1),
    n_la_Vatben = log(N_Vatben +1)
  ) %>%
  ungroup() %>%

  ## Visual-spatial functioning
  rowwise() %>%
  mutate(
    n_vp_NumLoc = log(N_NumLoc + 1),
    n_vp_FraLet = log(N_FraLet + 1)
  ) %>%
  ungroup()
  


data_final <- data_tmp %>%
  
  ### Memory
  mutate(
    n_me_WT_sum_z  = as.numeric(scale(n_me_WT_sum)),
    n_me_delayed_z  = as.numeric(scale(n_me_delayed)),
    n_me_recognition_z = as.numeric(scale(n_me_recognition)),
    n_me_vata_z  = as.numeric(scale(n_me_vata))
  ) %>%
  
  rowwise() %>%
  mutate(n_memory_1_z = mean(c_across(c(n_me_WT_sum_z, n_me_delayed_z, n_me_recognition_z)),na.rm = TRUE)) %>%
  
  mutate(n_memory_z = mean(c_across(cols = c(n_memory_1_z, n_me_vata_z)), na.rm = TRUE)) %>%
  
  mutate(na_count_memory = rowSums(is.na(across(c("n_me_WT_sum_z", "n_me_delayed_z", "n_me_recognition_z", "n_me_vata_z"))))) %>%
  ungroup() %>%
  
    
  ### Executive functioning
  mutate(
    n_ef_FL_sum_z      = as.numeric(scale(n_ef_FL_sum)),
    n_ef_TMT_index_z     = as.numeric(scale(n_ef_TMT_index)),
    n_ef_Strt_index_z = as.numeric(scale(n_ef_Strt_index))
  ) %>%
  
  rowwise() %>%
  mutate(n_executive_z = mean(c_across(cols = c(n_ef_FL_sum_z, n_ef_Strt_index_z)), na.rm = TRUE)) %>%
  mutate(na_count_executive = rowSums(is.na(across(c("n_ef_FL_sum_z", "n_ef_TMT_index_z", "n_ef_Strt_index_z"))))) %>%
  ungroup() %>%
  
  
  ### Attention
  mutate(
    n_ap_CRspv_z      = as.numeric(scale(n_ap_CRspv)),
    n_ap_TMTat_z     = as.numeric(scale(n_ap_TMTat)),
    n_ap_Strt1_z = as.numeric(scale(n_ap_Strt1)),
    n_ap_Strt2_z = as.numeric(scale(n_ap_Strt2))
  ) %>%
  
  rowwise() %>%
  mutate(n_attsp_z = mean(c_across(cols = c(n_ap_CRspv_z, n_ap_TMTat_z, n_ap_Strt1_z,n_ap_Strt2_z)), na.rm = TRUE)) %>%
  mutate(na_count_attention = rowSums(is.na(across(c("n_ap_CRspv_z", "n_ap_TMTat_z", "n_ap_Strt1_z","n_ap_Strt2_z"))))) %>%
  ungroup() %>%
  
  ### Language
  mutate(
    n_la_FL60_z      = as.numeric(scale(n_la_FL60)),
    n_la_Vatben_z     = as.numeric(scale(n_la_Vatben))
  ) %>%
  
  rowwise() %>%
  mutate(n_language_z = mean(c_across(cols = c(n_la_FL60_z, n_la_Vatben_z)), na.rm = TRUE)) %>%
  mutate(na_count_language = rowSums(is.na(across(c("n_la_FL60_z", "n_la_Vatben_z"))))) %>%
  ungroup() %>%
  
  ### Visualspatial functioning
  mutate(
    n_vp_NumLoc_z      = as.numeric(scale(n_vp_NumLoc)),
    n_vp_FraLet_z     = as.numeric(scale(n_vp_FraLet))
  ) %>%
  
  rowwise() %>%
  mutate(n_visualspatial_z = mean(c_across(cols = c(n_vp_NumLoc_z, n_vp_FraLet_z)), na.rm = TRUE)) %>%
  mutate(na_count_visualspatial = rowSums(is.na(across(c("n_vp_NumLoc_z", "n_vp_FraLet_z"))))) %>%
  ungroup()


C1_domains_Annebet <- data_final %>%
  mutate(Diagnosis_N = case_when(
    D_diagnosis == "SCD" & amyloid_status == "A-" ~ "SCD(A-)",
    D_diagnosis == "AD" & amyloid_status == "A+" ~ "AD(A+)",
    D_diagnosis == "FTD" & amyloid_status == "A-" ~ "FTD(A-)",
    D_diagnosis == "FTD" & amyloid_status == "A+" ~ "FTD(A+)",
    D_diagnosis == "DLB" & amyloid_status == "A-" ~ "DLB(A-)",
    D_diagnosis == "DLB" & amyloid_status == "A+" ~ "DLB(A+)",
  ))

save(C1_domains_Annebet,file = "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/data/C1_domains_Annebet_no_imputation.Rdata")



table(C1_domains_Annebet$D_diagnosis[!C1_domains_Annebet$APOE_status == "Missing"],C1_domains_Annebet$amyloid_status[!C1_domains_Annebet$APOE_status == "Missing"])


#view(C1_domains_Annebet[C1_domains_Annebet$Diagnosis_N == "DLB(A+)",c("n_memory_z","N_TMTAt","n_attsp_z","n_language_z")])
```

# S-Table 2: neuropsychological tests missing report
```{r}
load("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/data/C1_domains_Annebet_no_imputation.Rdata")

data_ADC_analysis = C1_domains_Annebet

data_ADC_analysis_cog <- data_ADC_analysis %>%
  mutate(Diagnosis_N = case_when(
    D_diagnosis == "SCD" & amyloid_status == "A-" ~ "SCD(A-)",
    D_diagnosis == "AD" & amyloid_status == "A+" ~ "AD(A+)",
    D_diagnosis == "FTD" & amyloid_status == "A-" ~ "FTD(A-)",
    D_diagnosis == "FTD" & amyloid_status == "A+" ~ "FTD(A+)",
    D_diagnosis == "DLB" & amyloid_status == "A-" ~ "DLB(A-)",
    D_diagnosis == "DLB" & amyloid_status == "A+" ~ "DLB(A+)",
  )) %>%
  mutate(Diagnosis_N = factor(Diagnosis_N,c("SCD(A-)","AD(A+)","FTD(A-)","FTD(A+)","DLB(A-)","DLB(A+)"))) %>%
  mutate(across(all_of(c("V_MMSE")), scale)) %>%
  filter(!I_ID == 3329)


missing_report(data_ADC_analysis_cog[,c("N_15WT1","N_15WT2","N_15WT3","N_15WT4","N_15WT5","N_15WTui","N_15WTvn","N_15WTvp","N_VATA1","N_VATA2")])
missing_report(data_ADC_analysis_cog[,c("N_FLD","N_FLA","N_FLT","N_Strt1","N_Strt2","N_Strt3")])
missing_report(data_ADC_analysis_cog[,c("N_CRspv","N_TMTAt","N_Strt1","N_Strt2")])


missing_report(data_ADC_analysis_cog[,c("n_vp_NumLoc_z","n_vp_FraLet_z")])
missing_report(data_ADC_analysis_cog[data_ADC_analysis_cog$D_diagnosis == "DLB",c("n_vp_NumLoc_z","n_vp_FraLet_z")])
```

# Figure 1: radar plot
```{r}
load("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/data/C1_domains_Annebet_no_imputation.Rdata")

data_ADC_analysis <- C1_domains_Annebet
table(data_ADC_analysis$D_diagnosis,data_ADC_analysis$amyloid_status)

data_ADC_analysis_cog <- data_ADC_analysis %>%
  mutate(Diagnosis_N = case_when(
    D_diagnosis == "SCD" & amyloid_status == "A-" ~ "SCD(A-)",
    D_diagnosis == "AD" & amyloid_status == "A+" ~ "AD(A+)",
    D_diagnosis == "FTD" & amyloid_status == "A-" ~ "FTD(A-)",
    D_diagnosis == "FTD" & amyloid_status == "A+" ~ "FTD(A+)",
    D_diagnosis == "DLB" & amyloid_status == "A-" ~ "DLB(A-)",
    D_diagnosis == "DLB" & amyloid_status == "A+" ~ "DLB(A+)",
  )) %>%
  mutate(Diagnosis_N = factor(Diagnosis_N,c("SCD(A-)","AD(A+)","FTD(A-)","FTD(A+)","DLB(A-)","DLB(A+)"))) %>%
  mutate(across(all_of(c("V_MMSE")), scale)) %>%
  filter(!I_ID == 3329)


data_N_A_analysis_AD <- data_ADC_analysis_cog %>% filter(D_diagnosis == "AD") 

# descrTable(Diagnosis_N ~amyloid_age+Sex+Edu+APOE_status+V_MMSE + n_memory_z + n_executive_z + n_attsp_z + n_language_z,
#           include.miss = TRUE,data_ADC_analysis_cog,max.ylev = 7,show.p.overall = FALSE,show.p.mul = FALSE)

# export2xls(table, "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/plots/plots_cognitive_domains/table.xlsx", which.table="descr", nmax = TRUE, nmax.method = 1)


################################ FTD & DLB ################################ 
data_FTD_minus <- data.frame(
  Global = c(1, -1, 0.09,-0.06),
  Memory = c(1, -1, 0.17,-0.15),
  Executive = c(1, -1, 0.34,-0.32),
  Attention = c(1, -1, -0.17,-0.28),
  Language = c(1, -1, -0.40,-0.65)
)
rownames(data_FTD_minus) <- c("Max", "Min","FTD(A-)","FTD(A+)")



data_DLB_minus <- data.frame(
  Global = c(1, -1, 0.02,-0.29),
  Memory = c(1, -1, 0.13,-0.10),
  Executive = c(1, -1, -0.55,-0.52),
  Attention = c(1, -1, -0.59,-0.74),
  Language = c(1, -1, 0.02,-0.27)
  
)
rownames(data_DLB_minus) <- c("Max", "Min","DLB(A-)","DLB(A+)")

library(fmsb)

create_beautiful_radarchart <- function(data, color = NULL,title = NULL){
  seg <- 4
  series_idx <- 3:nrow(data)
  series_colors <- rep_len(color, length(series_idx))
  theta <- seq(90, 450, length.out = ncol(data) + 1)[seq_len(ncol(data))] * pi / 180
  xx <- cos(theta)
  yy <- sin(theta)

  radarchart(
    data,
    axistype = 1,
    pty = 16,
    pcol = series_colors,
    pfcol = scales::alpha(series_colors, 0.12),
    plty = 1,
    plwd = 3.8,
    cglcol = "#B0B0B0",
    cglty = 1,
    cglwd = 1.8,
    axislabcol = "#2F2F2F",
    calcex = 1.2,
    vlabels = rep("", ncol(data)),
    vlcex = 1.3,
    caxislabels = c(-1, -0.5, 0, 0.5, 1),
    title = title
  )

  text(xx * 1.30, yy * 1.30, labels = colnames(data), cex = 1.3, font = 2, col = "#141414")

  for (i in seq_along(series_idx)) {
    scale <- 1 / (seg + 1) +
      (as.numeric(data[series_idx[i], ]) - as.numeric(data[2, ])) /
      (as.numeric(data[1, ]) - as.numeric(data[2, ])) * seg / (seg + 1)

    points(xx * scale, yy * scale, pch = 21, bg = "white", col = "white", cex = 3.2)
    points(xx * scale, yy * scale, pch = 21, bg = series_colors[i], col = series_colors[i], lwd = 1.4, cex = 2.2)
  }
}

draw_radar_legend <- function(labels, colors) {
  legend(
    x = "bottom",
    inset = c(0, -0.22),
    legend = labels,
    horiz = TRUE,
    bty = "n",
    pch = 21,
    pt.bg = colors,
    col = colors,
    text.col = "#1F1F1F",
    cex = 1.45,
    pt.cex = 2.6
  )
}

save_single_radar <- function(file, data, color) {
  pdf(file, width = 6.8, height = 6.0)
  op <- par(no.readonly = TRUE)
  par(mfrow = c(1, 1), mar = c(4.8, 1.1, 1.3, 1.1), xpd = NA)
  create_beautiful_radarchart(data, color = color)
  draw_radar_legend(rownames(data[-c(1, 2), ]), color)
  par(op)
  dev.off()
}


pdf("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/Figure1/FTD_DLB_radar_comparison.pdf", width = 11.6, height = 6.2)
op <- par(no.readonly = TRUE)
par(mfrow = c(1, 2), mar = c(4.8, 1.2, 1.3, 1.2), xpd = NA)

create_beautiful_radarchart(data_FTD_minus, color = c("#73929B", "#C06C84"))
draw_radar_legend(rownames(data_FTD_minus[-c(1, 2), ]), c("#73929B", "#C06C84"))

create_beautiful_radarchart(data_DLB_minus, color = c("#73929B", "#C06C84"))
draw_radar_legend(rownames(data_DLB_minus[-c(1, 2), ]), c("#73929B", "#C06C84"))

par(op)
dev.off()




################################ SCD groups ################################ 
table <- descrTable(D_diagnosis ~amyloid_age+Edu+APOE_status+V_MMSE + n_memory_z + n_executive_z + n_attsp_z + n_language_z,
          include.miss = TRUE,data_ADC_analysis_cog,max.ylev = 7,show.p.overall = FALSE,show.p.mul = FALSE)

#export2csv(table, "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/plots/plots_cognitive_domains/table.csv", which.table="descr", sep=",", nmax = TRUE, nmax.method = 1)


data_SCD <- data.frame(
  Global = c(1, -1, 0.83),
  Memory = c(1, -1, 0.74),
  Executive = c(1, -1, 0.47),
  Attention = c(1, -1, 0.59),
  Language = c(1, -1, 0.53)
)
rownames(data_SCD) <- c("Max", "Min","CU")

save_single_radar(
  "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/Figure1/radar_comparison_SCD.pdf",
  data_SCD,
  c("grey55")
)



data_AD <- data.frame(
  Global = c(1, -1, -0.53),
  Memory = c(1, -1, -0.58),
  Executive = c(1, -1, -0.36),
  Attention = c(1, -1, -0.42),
  Language = c(1, -1, -0.26)
)
rownames(data_AD) <- c("Max", "Min","AD")

save_single_radar(
  "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/Figure1/radar_comparison_AD.pdf",
  data_AD,
  c("grey55")
)




data_FTD <- data.frame(
  Global = c(1, -1, 0.06),
  Memory = c(1, -1, 0.11),
  Executive = c(1, -1, -0.34),
  Attention = c(1, -1, -0.19),
  Language = c(1, -1, -0.45)
)
rownames(data_FTD) <- c("Max", "Min","FTD")

save_single_radar(
  "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/Figure1/radar_comparison_FTD.pdf",
  data_FTD,
  c("grey55")
)


data_DLB <- data.frame(
  Global = c(1, -1, -0.15),
  Memory = c(1, -1, 0.01),
  Executive = c(1, -1, -0.53),
  Attention = c(1, -1, -0.67),
  Language = c(1, -1, -0.14)
)
rownames(data_DLB) <- c("Max", "Min","DLB")

save_single_radar(
  "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/Figure1/radar_comparison_DLB.pdf",
  data_DLB,
  c("grey55")
)



################################  check significance 1 ################################ 
library(bruceR)
data_ADC_test <- data_ADC_analysis_cog %>% 
  filter(D_diagnosis %in% c("SCD","DLB"))
# V_MMSE + n_memory_z + n_executive_z + n_attsp_z + n_language_z 


results = MASS::rlm(n_language_z ~ D_diagnosis+amyloid_age+Sex+Edu,data =data_ADC_test)
coef_table <- summary(results)$coefficients
results_df <- as.data.frame(coef_table)
colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
robust_test = sfsmisc::f.robftest(results,var = names(results$coefficients)[2])

p_two <- robust_test$p.value
coef_val <- results_df$Estimate[2]  # or however you extract the beta
std_val <- results_df$Std_Error[2]
# One-sided p-values in both directions
p_one_minus <- p_two / 2      # H1: effect < 0
p_one_plus  <- p_two / 2      # H1: effect > 0

# Keep the one in the "wrong" direction as 1
if (coef_val < 0) {
    p_one_plus <- 1
} else if (coef_val > 0) {
    p_one_minus <- 1
}
paste(coef_val, std_val, p_one_minus, p_one_plus)




################################  check significance ################################ 
library(bruceR)
data_FTD <- data_ADC_analysis_cog %>% 
  filter(Diagnosis_N %in% c("DLB(A-)","DLB(A+)")) %>% 
  mutate(Diagnosis_N  = factor(Diagnosis_N, c("DLB(A-)","DLB(A+)")))
# V_MMSE + n_memory_z + n_executive_z + n_attsp_z + n_language_z 


results = MASS::rlm(n_language_z ~ Diagnosis_N+amyloid_age+Sex+Edu,data =data_FTD)
coef_table <- summary(results)$coefficients
results_df <- as.data.frame(coef_table)
colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
robust_test = sfsmisc::f.robftest(results,var = names(results$coefficients)[2])

p_two <- robust_test$p.value
coef_val <- results_df$Estimate[2]  # or however you extract the beta
std_val <- results_df$Std_Error[2]
# One-sided p-values in both directions
p_one_minus <- p_two / 2      # H1: effect < 0
p_one_plus  <- p_two / 2      # H1: effect > 0

# Keep the one in the "wrong" direction as 1
if (coef_val < 0) {
    p_one_plus <- 1
} else if (coef_val > 0) {
    p_one_minus <- 1
}
paste(coef_val, std_val, p_one_minus, p_one_plus)
```

# S-Figure 5: radar plot with spatial functioning
```{r}
load("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/data/C1_domains_Annebet_no_imputation.Rdata")

data_ADC_analysis <- C1_domains_Annebet
table(data_ADC_analysis$D_diagnosis,data_ADC_analysis$amyloid_status)

data_ADC_analysis_cog <- data_ADC_analysis %>%
  mutate(Diagnosis_N = case_when(
    D_diagnosis == "SCD" & amyloid_status == "A-" ~ "SCD(A-)",
    D_diagnosis == "AD" & amyloid_status == "A+" ~ "AD(A+)",
    D_diagnosis == "FTD" & amyloid_status == "A-" ~ "FTD(A-)",
    D_diagnosis == "FTD" & amyloid_status == "A+" ~ "FTD(A+)",
    D_diagnosis == "DLB" & amyloid_status == "A-" ~ "DLB(A-)",
    D_diagnosis == "DLB" & amyloid_status == "A+" ~ "DLB(A+)",
  )) %>%
  mutate(Diagnosis_N = factor(Diagnosis_N,c("SCD(A-)","AD(A+)","FTD(A-)","FTD(A+)","DLB(A-)","DLB(A+)"))) %>%
  mutate(across(all_of(c("V_MMSE")), scale)) %>%
  filter(!I_ID == 3329)


data_N_A_analysis_AD <- data_ADC_analysis_cog %>% filter(D_diagnosis == "AD") 

table <- descrTable(Diagnosis_N ~amyloid_age+Sex+Edu+APOE_status+V_MMSE + n_memory_z + n_executive_z + n_attsp_z + n_language_z+n_visualspatial_z,
          include.miss = TRUE,data_ADC_analysis_cog,max.ylev = 7,show.p.overall = FALSE,show.p.mul = FALSE)


descrTable(D_diagnosis ~amyloid_age+Sex+Edu+APOE_status+V_MMSE + n_memory_z + n_executive_z + n_attsp_z + n_language_z+n_visualspatial_z,
          include.miss = TRUE,data_ADC_analysis_cog,max.ylev = 7,show.p.overall = FALSE,show.p.mul = FALSE)

# export2xls(table, "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/plots/plots_cognitive_domains/table.xlsx", which.table="descr", nmax = TRUE, nmax.method = 1)



################################ DLB(A-) & DLB(A+) ################################ 
data_DLB_minus <- data.frame(
  Global = c(1, -1, 0.02,-0.29),
  Memory = c(1, -1, 0.13,-0.10),
  Executive = c(1, -1, -0.55,-0.52),
  Attention = c(1, -1, -0.59,-0.74),
  Language = c(1, -1, 0.02,-0.27),
  Visual_spatial = c(1, -1, -0.22, -0.38)
  
)
rownames(data_DLB_minus) <- c("Max", "Min","DLB(A-)","DLB(A+)")

library(fmsb)
create_beautiful_radarchart <- function(data, color = NULL,title = NULL){
  seg <- 4
  series_idx <- 3:nrow(data)
  series_colors <- rep_len(color, length(series_idx))
  theta <- seq(90, 450, length.out = ncol(data) + 1)[seq_len(ncol(data))] * pi / 180
  xx <- cos(theta)
  yy <- sin(theta)

  radarchart(
    data,
    axistype = 1,
    pty = 16,
    pcol = series_colors,
    pfcol = scales::alpha(series_colors, 0.12),
    plty = 1,
    plwd = 3.8,
    cglcol = "#B0B0B0",
    cglty = 1,
    cglwd = 1.8,
    axislabcol = "#2F2F2F",
    calcex = 1.2,
    vlabels = rep("", ncol(data)),
    vlcex = 1.3,
    caxislabels = c(-1, -0.5, 0, 0.5, 1),
    title = title
  )

  text(xx * 1.30, yy * 1.30, labels = colnames(data), cex = 1.3, font = 2, col = "#141414")

  for (i in seq_along(series_idx)) {
    scale <- 1 / (seg + 1) +
      (as.numeric(data[series_idx[i], ]) - as.numeric(data[2, ])) /
      (as.numeric(data[1, ]) - as.numeric(data[2, ])) * seg / (seg + 1)

    points(xx * scale, yy * scale, pch = 21, bg = "white", col = "white", cex = 3.2)
    points(xx * scale, yy * scale, pch = 21, bg = series_colors[i], col = series_colors[i], lwd = 1.4, cex = 2.2)
  }
}

draw_radar_legend <- function(labels, colors) {
  legend(
    x = "bottom",
    inset = c(0, -0.22),
    legend = labels,
    horiz = TRUE,
    bty = "n",
    pch = 21,
    pt.bg = colors,
    col = colors,
    text.col = "#1F1F1F",
    cex = 1.45,
    pt.cex = 2.6
  )
}

save_single_radar <- function(file, data, color) {
  pdf(file, width = 6.8, height = 6.0)
  op <- par(no.readonly = TRUE)
  par(mfrow = c(1, 1), mar = c(4.8, 1.1, 1.3, 1.1), xpd = NA)
  create_beautiful_radarchart(data, color = color)
  draw_radar_legend(rownames(data[-c(1, 2), ]), color)
  par(op)
  dev.off()
}


save_single_radar(
  "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure5/radar_comparison_DLB_A_status.pdf",
  data_DLB_minus,
  c("#73929B", "#C06C84")
)


################################ SCD groups ################################ 
table <- descrTable(D_diagnosis ~amyloid_age+Edu+APOE_status+V_MMSE + n_memory_z + n_executive_z + n_attsp_z + n_language_z,
          include.miss = TRUE,data_ADC_analysis_cog,max.ylev = 7,show.p.overall = FALSE,show.p.mul = FALSE)

data_SCD <- data.frame(
  Global = c(1, -1, 0.83),
  Memory = c(1, -1, 0.74),
  Executive = c(1, -1, 0.47),
  Attention = c(1, -1, 0.59),
  Language = c(1, -1, 0.53),
  Visual_spatial = c(1, -1, 0.40)
)
rownames(data_SCD) <- c("Max", "Min","CU")

save_single_radar(
  "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure5/radar_comparison_SCD.pdf",
  data_SCD,
  c("grey55")
)


data_DLB <- data.frame(
  Global = c(1, -1, -0.15),
  Memory = c(1, -1, 0.01),
  Executive = c(1, -1, -0.53),
  Attention = c(1, -1, -0.67),
  Language = c(1, -1, -0.14),
  Visual_spatial = c(1, -1, -0.31)
)
rownames(data_DLB) <- c("Max", "Min","DLB")

save_single_radar(
  "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure5/radar_comparison_DLB.pdf",
  data_DLB,
  c("grey55")
)




################################  check significance 1 ################################ 
library(bruceR)
data_ADC_test <- data_ADC_analysis_cog %>% 
  filter(D_diagnosis %in% c("SCD","DLB"))
# V_MMSE + n_memory_z + n_executive_z + n_attsp_z + n_language_z 


results = MASS::rlm(n_visualspatial_z ~ D_diagnosis+amyloid_age+Sex+Edu,data =data_ADC_test)
coef_table <- summary(results)$coefficients
results_df <- as.data.frame(coef_table)
colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
robust_test = sfsmisc::f.robftest(results,var = names(results$coefficients)[2])

p_two <- robust_test$p.value
coef_val <- results_df$Estimate[2]  # or however you extract the beta
std_val <- results_df$Std_Error[2]
# One-sided p-values in both directions
p_one_minus <- p_two / 2      # H1: effect < 0
p_one_plus  <- p_two / 2      # H1: effect > 0

# Keep the one in the "wrong" direction as 1
if (coef_val < 0) {
    p_one_plus <- 1
} else if (coef_val > 0) {
    p_one_minus <- 1
}
paste(coef_val, std_val, p_one_minus, p_one_plus)




################################  check significance ################################ 
library(bruceR)
data_FTD <- data_ADC_analysis_cog %>% 
  filter(Diagnosis_N %in% c("DLB(A-)","DLB(A+)")) %>% 
  mutate(Diagnosis_N  = factor(Diagnosis_N, c("DLB(A-)","DLB(A+)")))
# V_MMSE + n_memory_z + n_executive_z + n_attsp_z + n_language_z 


results = MASS::rlm(n_language_z ~ Diagnosis_N+amyloid_age+Sex+Edu,data =data_FTD)
coef_table <- summary(results)$coefficients
results_df <- as.data.frame(coef_table)
colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
robust_test = sfsmisc::f.robftest(results,var = names(results$coefficients)[2])

p_two <- robust_test$p.value
coef_val <- results_df$Estimate[2]  # or however you extract the beta
std_val <- results_df$Std_Error[2]
# One-sided p-values in both directions
p_one_minus <- p_two / 2      # H1: effect < 0
p_one_plus  <- p_two / 2      # H1: effect > 0

# Keep the one in the "wrong" direction as 1
if (coef_val < 0) {
    p_one_plus <- 1
} else if (coef_val > 0) {
    p_one_minus <- 1
}
paste(coef_val, std_val, p_one_minus, p_one_plus)
```


# S-Figure 9: APOE corrected radar plot
```{r}
load("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/data/C1_domains_Annebet_no_imputation.Rdata")

data_ADC_analysis <- C1_domains_Annebet
table(data_ADC_analysis$D_diagnosis,data_ADC_analysis$amyloid_status)

data_ADC_analysis_cog <- data_ADC_analysis %>%
  mutate(Diagnosis_N = case_when(
    D_diagnosis == "SCD" & amyloid_status == "A-" ~ "SCD(A-)",
    D_diagnosis == "AD" & amyloid_status == "A+" ~ "AD(A+)",
    D_diagnosis == "FTD" & amyloid_status == "A-" ~ "FTD(A-)",
    D_diagnosis == "FTD" & amyloid_status == "A+" ~ "FTD(A+)",
    D_diagnosis == "DLB" & amyloid_status == "A-" ~ "DLB(A-)",
    D_diagnosis == "DLB" & amyloid_status == "A+" ~ "DLB(A+)",
  )) %>%
  mutate(Diagnosis_N = factor(Diagnosis_N,c("SCD(A-)","AD(A+)","FTD(A-)","FTD(A+)","DLB(A-)","DLB(A+)"))) %>%
  mutate(across(all_of(c("V_MMSE")), scale)) %>%
  filter(!I_ID == 3329)


data_N_A_analysis_AD <- data_ADC_analysis_cog %>% filter(D_diagnosis == "AD") 

table <- descrTable(Diagnosis_N ~amyloid_age+Sex+Edu+APOE_status+V_MMSE + n_memory_z + n_executive_z + n_attsp_z + n_language_z+n_visualspatial_z,
          include.miss = TRUE,data_ADC_analysis_cog,max.ylev = 7,show.p.overall = FALSE,show.p.mul = FALSE)


descrTable(D_diagnosis ~amyloid_age+Sex+Edu+APOE_status+V_MMSE + n_memory_z + n_executive_z + n_attsp_z + n_language_z+n_visualspatial_z,
          include.miss = TRUE,data_ADC_analysis_cog,max.ylev = 7,show.p.overall = FALSE,show.p.mul = FALSE)

# export2xls(table, "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/plots/plots_cognitive_domains/table.xlsx", which.table="descr", nmax = TRUE, nmax.method = 1)



################################ DLB(A-) & DLB(A+) ################################ 
data_DLB_minus <- data.frame(
  Global = c(1, -1, 0.02,-0.29),
  Memory = c(1, -1, 0.13,-0.10),
  Executive = c(1, -1, -0.55,-0.52),
  Attention = c(1, -1, -0.59,-0.74),
  Language = c(1, -1, 0.02,-0.27),
  Visual_spatial = c(1, -1, -0.22, -0.38)
  
)
rownames(data_DLB_minus) <- c("Max", "Min","DLB(A-)","DLB(A+)")

library(fmsb)
create_beautiful_radarchart <- function(data, color = NULL,title = NULL){
  seg <- 4
  series_idx <- 3:nrow(data)
  series_colors <- rep_len(color, length(series_idx))
  theta <- seq(90, 450, length.out = ncol(data) + 1)[seq_len(ncol(data))] * pi / 180
  xx <- cos(theta)
  yy <- sin(theta)

  radarchart(
    data,
    axistype = 1,
    pty = 16,
    pcol = series_colors,
    pfcol = scales::alpha(series_colors, 0.12),
    plty = 1,
    plwd = 3.8,
    cglcol = "#B0B0B0",
    cglty = 1,
    cglwd = 1.8,
    axislabcol = "#2F2F2F",
    calcex = 1.2,
    vlabels = rep("", ncol(data)),
    vlcex = 1.3,
    caxislabels = c(-1, -0.5, 0, 0.5, 1),
    title = title
  )

  text(xx * 1.30, yy * 1.30, labels = colnames(data), cex = 1.3, font = 2, col = "#141414")

  for (i in seq_along(series_idx)) {
    scale <- 1 / (seg + 1) +
      (as.numeric(data[series_idx[i], ]) - as.numeric(data[2, ])) /
      (as.numeric(data[1, ]) - as.numeric(data[2, ])) * seg / (seg + 1)

    points(xx * scale, yy * scale, pch = 21, bg = "white", col = "white", cex = 3.2)
    points(xx * scale, yy * scale, pch = 21, bg = series_colors[i], col = series_colors[i], lwd = 1.4, cex = 2.2)
  }
}

draw_radar_legend <- function(labels, colors) {
  legend(
    x = "bottom",
    inset = c(0, -0.22),
    legend = labels,
    horiz = TRUE,
    bty = "n",
    pch = 21,
    pt.bg = colors,
    col = colors,
    text.col = "#1F1F1F",
    cex = 1.45,
    pt.cex = 2.6
  )
}

save_single_radar <- function(file, data, color) {
  pdf(file, width = 6.8, height = 6.0)
  op <- par(no.readonly = TRUE)
  par(mfrow = c(1, 1), mar = c(4.8, 1.1, 1.3, 1.1), xpd = NA)
  create_beautiful_radarchart(data, color = color)
  draw_radar_legend(rownames(data[-c(1, 2), ]), color)
  par(op)
  dev.off()
}


save_single_radar(
  "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure5/radar_comparison_DLB_A_status.pdf",
  data_DLB_minus,
  c("#73929B", "#C06C84")
)


################################  check significance ################################ 
library(bruceR)
data_FTD <- data_ADC_analysis_cog %>% 
  filter(Diagnosis_N %in% c("DLB(A-)","DLB(A+)")) %>% 
  mutate(Diagnosis_N  = factor(Diagnosis_N, c("DLB(A-)","DLB(A+)")))
# V_MMSE + n_memory_z + n_executive_z + n_attsp_z + n_language_z 


results = MASS::rlm(n_language_z ~ Diagnosis_N+amyloid_age+Sex+Edu+APOE_status,data =data_FTD)
coef_table <- summary(results)$coefficients
results_df <- as.data.frame(coef_table)
colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
robust_test = sfsmisc::f.robftest(results,var = names(results$coefficients)[2])

p_two <- robust_test$p.value
coef_val <- results_df$Estimate[2]  # or however you extract the beta
std_val <- results_df$Std_Error[2]
# One-sided p-values in both directions
p_one_minus <- p_two / 2      # H1: effect < 0
p_one_plus  <- p_two / 2      # H1: effect > 0

# Keep the one in the "wrong" direction as 1
if (coef_val < 0) {
    p_one_plus <- 1
} else if (coef_val > 0) {
    p_one_minus <- 1
}
paste(coef_val, std_val, p_one_minus, p_one_plus)

```


```{r}
# Figure 1: radar plot

library(bruceR)
data_ADC_test <- data_ADC_analysis_cog %>% 
  filter(D_diagnosis %in% c("SCD","FTD"))
# V_MMSE + n_memory_z + n_executive_z + n_attsp_z + n_language_z 


results = MASS::rlm(V_MMSE ~ D_diagnosis+amyloid_age+Sex+Edu,data =data_ADC_test)
coef_table <- summary(results)$coefficients
results_df <- as.data.frame(coef_table)
colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
robust_test = sfsmisc::f.robftest(results,var = names(results$coefficients)[2])

p_two <- robust_test$p.value
coef_val <- results_df$Estimate[2]  # or however you extract the beta
std_val <- results_df$Std_Error[2]
# One-sided p-values in both directions
p_one_minus <- p_two / 2      # H1: effect < 0
p_one_plus  <- p_two / 2      # H1: effect > 0

# Keep the one in the "wrong" direction as 1
if (coef_val < 0) {
    p_one_plus <- 1
} else if (coef_val > 0) {
    p_one_minus <- 1
}
paste(coef_val, std_val, p_one_minus, p_one_plus)
```



# Table1: average grey matter volume; average subcortical volume
```{r}
load("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/data/C1_domains_Annebet_no_imputation.Rdata")

names_Vol <- c(names(Freesurfer_output_718)[240:273],names(Freesurfer_output_718)[276:309])
names_Sub_2<- c("lh_Thalamus_volume","lh_Caudate_volume","lh_Putamen_volume","lh_Pallidum_volume","lh_Hippocampus_volume","lh_Amygdala_volume",
              "lh_Accumbens_area_volume","lh_choroid_plexus_volume",
              "rh_Thalamus_volume","rh_Caudate_volume","rh_Putamen_volume","rh_Pallidum_volume","rh_Hippocampus_volume","rh_Amygdala_volume",
              "rh_Accumbens_area_volume","rh_choroid_plexus_volume")


data_ADC_analysis <- C1_domains_Annebet
table(data_ADC_analysis$D_diagnosis,data_ADC_analysis$amyloid_status)

data_ADC_analysis_cog <- data_ADC_analysis %>%
  mutate(Diagnosis_N = case_when(
    D_diagnosis == "SCD" & amyloid_status == "A-" ~ "SCD(A-)",
    D_diagnosis == "AD" & amyloid_status == "A+" ~ "AD(A+)",
    D_diagnosis == "FTD" & amyloid_status == "A-" ~ "FTD(A-)",
    D_diagnosis == "FTD" & amyloid_status == "A+" ~ "FTD(A+)",
    D_diagnosis == "DLB" & amyloid_status == "A-" ~ "DLB(A-)",
    D_diagnosis == "DLB" & amyloid_status == "A+" ~ "DLB(A+)",
  )) %>%
  mutate(Diagnosis_N = factor(Diagnosis_N,c("SCD(A-)","AD(A+)","FTD(A-)","FTD(A+)","DLB(A-)","DLB(A+)"))) %>%
  mutate(across(all_of(c("V_MMSE")), scale)) %>%
  filter(!I_ID == 3329) %>%
  rowwise() %>%
  mutate(
    mean_gm_volume = mean(c_across(all_of(names_Vol)), na.rm = TRUE),
    mean_subcortical_volume = mean(c_across(all_of(names_Sub_2)), na.rm = TRUE)) %>%
  ungroup()

data_ADC_analysis_tmp <- data_ADC_analysis_cog %>% filter(Diagnosis_N %in% c("DLB(A-)","DLB(A+)"))
descrTable(Diagnosis_N ~amyloid_age+Sex+Edu+mean_gm_volume + mean_subcortical_volume,
          include.miss = TRUE,data_ADC_analysis_tmp,max.ylev = 7,show.p.overall = TRUE)



data_test <- data_ADC_analysis_cog %>%
  filter(Diagnosis_N %in% c("FTD(A-)","FTD(A+)")) %>%
  mutate(Diagnosis_N = factor(Diagnosis_N,c("FTD(A-)","FTD(A+)")))

re_post <- data_test %>% 
  MANOVA(dv="mean_gm_volume", between="Diagnosis_N",covariate = c("Sex","Edu")) %>%
  EMMEANS(effect = "Diagnosis_N",
  by = NULL,
  contrast = "pairwise",
  reverse = TRUE,
  p.adjust = "none",
  sd.pooled = NULL,
  model.type = "multivariate",
  digits = 3)

re_post$EMMEANS
```



# Supplementary Figure 1
```{r}
################## Visualization:  Supplementary plot Figure1 ################## 
data_ADC_vis <- data_ADC_analysis %>% filter(D_diagnosis %in% c("SCD","AD","DLB","FTD")) %>% 
  select(diff_L_M,diff_L_N,diff_L_D,diff_L_V) %>%
  pivot_longer(
    cols = everything(),
    names_to = "task",
    values_to = "score"
  ) %>%
  rename(Modalities = task,Days = score)

plot_vis_time <- ggplot(data_ADC_vis, aes(x = Days, fill = Modalities)) +
  geom_density(alpha = 0.5) +
  coord_cartesian(xlim = c(-30, 30)) +  # adjust range as needed
  #labs(title = "Time Interval", x = "Time Interval", y = "Density") +
  theme_minimal() +
  theme(text = element_text(size = 14), legend.position = "top")
#ggsave("/Users/jiangxiaofan/Desktop/projects/project_ADC_mega_copathology/plots/Figure0_time_interval.pdf",plot_vis_time,height = 8,width = 12,dpi = 300)

```



# Supplementary Figure 2: FTD subtypes
```{r}
# Create a sample data frame
data <- data.frame(
  category = c("A", "B", "C", "D"),
  value = c(10, 25, 15, 50)
)
 
# Add a column for the proportion
data$proportion <- data$value / sum(data$value)
 
# Add a column for the label position
data$label_pos <- cumsum(data$proportion) - data$proportion / 2
 
# Create the pie chart
p <- ggplot(data, aes(x = "", y = proportion, fill = category)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar("y", start = 0) +
  theme_void() + # Use a minimalist theme
  theme(legend.position = "right") + # Place the legend on the right
  geom_text(aes(y = label_pos, label = paste0(round(proportion * 100, 1), "%")), size = 5) + # Add percentage labels
  labs(title = "Sample Pie Chart", fill = "Category") # Add a title and label for the legend

```



```{r}
load("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/data/C1_domains_Annebet_no_imputation.Rdata")
data_ADC_analysis <- C1_domains_Annebet

library(dplyr)
library(stringr)

data = read_xlsx("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/data/subtypes_FTD_sterre_jcy.xlsx")

data$Subtype = data$subtype
unique(data$subtype)

# df should contain a column named `Subtype`
df <- data %>%
  mutate(
    # clean & normalize labels
    subtype_clean = Subtype |> 
      str_squish() |>                       # trim & collapse spaces
      str_replace_all("\\.+$", "") |>       # drop trailing dots (e.g., "Phenocopy FTD.")
      str_replace_all("\\s*\\+\\s*", "+"),  # normalize " + " to "+"
    subtype_lc = str_to_lower(subtype_clean),
    
    # main 5-way grouping
    ftd_group = case_when(
      # bvFTD spectrum (incl. phenocopy, bvAD, bvFTD+PSP)
      subtype_lc %in% c("possible bvftd","probable bvftd","definite bvftd",
                        "phenocopy bvftd","phenocopy ftd","bvftd phenotype",
                        "bvftd+psp","bvad")                                  ~ "bvFTD spectrum",
      # PPA spectrum
      subtype_lc %in% c("svppa","ass","nfvppa","ppa","ppa nno","svppa+vad")   ~ "PPA spectrum",
      # Right temporal variant
      subtype_lc == "rtvftd"                                                  ~ "Right temporal variant FTD",
      # Motor/parkinsonian overlap
      subtype_lc %in% c("ftd-als","psp","ppd","pd")                           ~ "Motor/parkinsonian overlap",
      # Other / Unclassified / Mimics
      subtype_lc %in% c("ftd nno","fns","psychiatrie","uitgesteld")           ~ "Other/Unclassified",
      
      TRUE                                                                     ~ "Other/Unclassified"
    ),
    
    ftd_group_Lianne_Sterre = case_when(
      # bvFTD spectrum (incl. phenocopy, bvAD, bvFTD+PSP)
      subtype_lc %in% c("possible bvftd","probable bvftd","definite bvftd","bvftd+psp") ~ "bvFTD spectrum",
      # PPA spectrum
      subtype_lc %in% c("svppa","ass","nfvppa","ppa","ppa nno","svppa+vad")   ~ "PPA spectrum",
      # Right temporal variant
      subtype_lc == "rtvftd"                                                  ~ "Right temporal variant FTD",
      # Motor/parkinsonian overlap
      subtype_lc %in% c("FTD-ALS")                           ~ "FTD-ALS",
      subtype_lc %in% c("PSP")                           ~ "PSP",
      TRUE                                          ~ "Other/Unclassified"
    )
  )



df2 <- data_ADC_analysis %>% filter(D_diagnosis == "FTD") %>% 
  select(I_ID,D_diagnosis,amyloid_status) %>% 
  as.data.frame() %>% mutate(I_ID = as.numeric(I_ID)) %>%
  left_join(as.data.frame(df),by = "I_ID")


table(df2$amyloid_status,df2$ftd_group)

table(df2$amyloid_status,df2$ftd_group_Lianne_Sterre)

fwrite(df2,"/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/data/subtypes_FTD_jcy.txt",sep = "\t")


data <- as.data.frame(table(df2$ftd_group[df2$amyloid_status == "A-"])) %>% mutate(category = Var1, value = Freq)



data$proportion <- data$value / sum(data$value)
data$label_pos <- cumsum(data$proportion) - data$proportion / 2
p1 <- ggplot(data, aes(x = "", y = proportion, fill = category)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar("y", start = 0) +
  theme_void() + # Use a minimalist theme
  theme(legend.position = "left")  # Place the legend on the right
  #labs(title = "Sample Pie Chart", fill = "Category") # Add a title and label for the legend
p1
ggsave("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/plots/plots_flow_chart/Figure_S2_FTD_2.pdf",p1,width = 5,height = 5)


table(df2$ftd_grou)
prop.table(table(df2$ftd_group))

prop.table(table(df2$ftd_group[df2$amyloid_status == "A-"]))
proportions <- prop.table(table(df2$ftd_group[df2$amyloid_status == "A+"]))

# Round the proportions to one decimal place and multiply by 100 to get percentages
rounded_percentages <- round(proportions * 100, 1)
```
