---
title: "project_ADC_mega_main"
---

```{r,include=F}
rm(list = ls())
library(readxl)
library(tidyverse)
library(ggseg)
library(compareGroups)
library(ggpubr)
library(compareGroups)
library(sjPlot)
library(naniar)
library(bruceR)
library(mice)
library(readxl)
library(ggh4x)
```

# correlation of two pipelines
# domains - domains
# pipeline: which one to choose


```{r,include=FALSE}
load("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/data/C1_domains_Annebet_no_imputation.Rdata")
data_ADC_analysis <- C1_domains_Annebet
table(data_ADC_analysis$D_diagnosis,data_ADC_analysis$amyloid_status)

data_ADC_analysis_cog <- data_ADC_analysis %>%
  mutate(Diagnosis_N = case_when(
    D_diagnosis == "SCD" & amyloid_status == "A-" ~ "SCD(A-)",
    D_diagnosis == "AD" & amyloid_status == "A+" ~ "AD(A+)",
    D_diagnosis == "FTD" & amyloid_status == "A-" ~ "FTD(A-)",
    D_diagnosis == "FTD" & amyloid_status == "A+" ~ "FTD(A+)",
    D_diagnosis == "DLB" & amyloid_status == "A-" ~ "DLB(A-)",
    D_diagnosis == "DLB" & amyloid_status == "A+" ~ "DLB(A+)",
  )) %>%
  mutate(Diagnosis_N = factor(Diagnosis_N,c("SCD(A-)","AD(A+)","FTD(A-)","FTD(A+)","DLB(A-)","DLB(A+)"))) %>%
  mutate(across(all_of(c("V_MMSE")), scale))

data_N_A_analysis_AD <- data_ADC_analysis_cog %>% filter(D_diagnosis == "AD")

descrTable(Diagnosis_N ~amyloid_age+Sex+Edu+APOE_status+V_MMSE + n_executive_z + n_attsp_z + n_language_z+lh_Putamen_volume+rh_Thalamus_volume+lh_Hippocampus_volume+rh_parsorbitalis_thickness,
include.miss = TRUE,data_ADC_analysis_cog,max.ylev = 7,show.p.overall = FALSE)
```


# functions loading
```{r}
theme_common <- function() {
  theme(
    axis.ticks.x = element_blank(),        
    axis.ticks.y = element_blank(),        
    panel.border = element_rect(fill = NA, color = "grey70", size = 2, linetype = "solid"),
    axis.text.x = element_text(size = 16, angle = 90, hjust = 0),
    axis.text.y = element_text(size = 16),
    legend.text = element_text(size = 20),
    legend.title = element_text(size = 16),
    legend.key.size = unit(1.5, "cm"),
    legend.box.spacing = unit(0.5, "cm"))
}



analysis_glm_cognition_case_control_interaction_volume_robust <- function(stats_freesurfer_PRS,cognition_domains){
    
    list=c("bankssts","caudal anterior cingulate","caudal middle frontal",
        "cuneus","entorhinal","frontal pole","fusiform","inferior parietal",
        "inferior temporal","insula","isthmus cingulate","lateral occipital",
        "lateral orbitofrontal","lingual","medial orbitofrontal","middle temporal",
        "paracentral","parahippocampal","pars opercularis","pars orbitalis",
        "pars triangularis","pericalcarine","postcentral","posterior cingulate",
        "precentral","precuneus","rostral anterior cingulate","rostral middle frontal",
        "superior frontal","superior parietal","superior temporal","supramarginal",
        "temporal pole","transverse temporal","thalamus proper","caudate","putamen",
        "pallidum","hippocampus","amygdala","accumbens area")
    
    list_sub = c("Thalamus_volume_mean","Caudate_volume_mean","Putamen_volume_mean",
                 "Pallidum_volume_mean","Hippocampus_volume_mean","Amygdala_volume_mean",
                 "Accumbens_area_volume_mean")
    CT_names <- c(str_c(str_replace_all(list[1:34]," ",""),"_volume_mean"),list_sub)
    
    DTK_Labes_For_Plot = list
    plot_data_L = as.data.frame(DTK_Labes_For_Plot)
    colnames(plot_data_L)[1] = "region"
    plot_data = plot_data_L
    
    plot_data$area_label = CT_names
    plot_data$main_Effect_ref=NA
    plot_data$main_pvalues_ref=NA
    plot_data$main_Effect_case=NA
    plot_data$main_pvalues_case=NA
    plot_data$interaction_Effect_case_ref=NA
    plot_data$interaction_Std_case_ref=NA
    plot_data$interaction_pvalues_case_ref=NA   
    attach(plot_data)
    
    for (im in 1:34) {
      Region_AV1451_v1 = plot_data$area_label[im]
      funtion_glm = as.formula(paste0(cognition_domains,"~",Region_AV1451_v1,"+Diagnosis_N+",Region_AV1451_v1,":Diagnosis_N+amyloid_age+Sex+Edu+eTIV"))
      results = MASS::rlm(funtion_glm,data =stats_freesurfer_PRS)
  
      coef_table <- summary(results)$coefficients
      results_df <- as.data.frame(coef_table)
      colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
      results_df <- results_df %>% mutate(p_value = 2 * (1 - pnorm(abs(t_value))))
      p_values = sfsmisc::f.robftest(results,var = names(results$coefficients)[8])

      plot_data$interaction_Effect_case_ref[im] = results_df[8,1]
      plot_data$interaction_Std_case_ref[im] = results_df[8,2]
      plot_data$interaction_pvalues_case_ref[im] = p_values$p.value
    }
    
    for (im in 35:41) {
      Region_AV1451_v1 = plot_data$area_label[im]
      Region_AV1451_ALL = plot_data$area_label
      funtion_glm = as.formula(paste0(cognition_domains,"~",Region_AV1451_v1,"+Diagnosis_N+",Region_AV1451_v1,
                                      ":Diagnosis_N+amyloid_age+Sex+Edu+eTIV"))
      results = MASS::rlm(funtion_glm,data =stats_freesurfer_PRS)
      
      coef_table <- summary(results)$coefficients
      results_df <- as.data.frame(coef_table)
      colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
      results_df <- results_df %>% mutate(p_value = 2 * (1 - pnorm(abs(t_value))))
      p_values = sfsmisc::f.robftest(results,var = names(results$coefficients)[8])
      
      plot_data$interaction_Effect_case_ref[im] = results_df[8,1]
      plot_data$interaction_Std_case_ref[im] = results_df[8,2]
      plot_data$interaction_pvalues_case_ref[im] = p_values$p.value
      
    }
    plot_data <- plot_data %>% 
      dplyr::select(region,area_label,interaction_Effect_case_ref,interaction_Std_case_ref,interaction_pvalues_case_ref) %>%
      distinct(area_label,.keep_all = TRUE) %>%
      mutate(cognition = cognition_domains)
    
    return(plot_data)
}


analysis_glm_cognition_case_control_interaction_thickness_robust <- function(stats_freesurfer_PRS,cognition_domains){
    
    list=c("bankssts","caudal anterior cingulate","caudal middle frontal",
        "cuneus","entorhinal","frontal pole","fusiform","inferior parietal",
        "inferior temporal","insula","isthmus cingulate","lateral occipital",
        "lateral orbitofrontal","lingual","medial orbitofrontal","middle temporal",
        "paracentral","parahippocampal","pars opercularis","pars orbitalis",
        "pars triangularis","pericalcarine","postcentral","posterior cingulate",
        "precentral","precuneus","rostral anterior cingulate","rostral middle frontal",
        "superior frontal","superior parietal","superior temporal","supramarginal",
        "temporal pole","transverse temporal","thalamus proper","caudate","putamen",
        "pallidum","hippocampus","amygdala","accumbens area")
    
    list_sub = c("Thalamus_volume_mean","Caudate_volume_mean","Putamen_volume_mean",
                 "Pallidum_volume_mean","Hippocampus_volume_mean","Amygdala_volume_mean",
                 "Accumbens_area_volume_mean")
    CT_names <- c(str_c(str_replace_all(list[1:34]," ",""),"_thickness_mean"),list_sub)
    
    DTK_Labes_For_Plot = list
    plot_data_L = as.data.frame(DTK_Labes_For_Plot)
    colnames(plot_data_L)[1] = "region"
    plot_data = plot_data_L
    
    plot_data$area_label = CT_names
    plot_data$main_Effect_ref=NA
    plot_data$main_pvalues_ref=NA
    plot_data$main_Effect_case=NA
    plot_data$main_pvalues_case=NA
    plot_data$interaction_Effect_case_ref=NA
    plot_data$interaction_Std_case_ref=NA
    plot_data$interaction_pvalues_case_ref=NA   
    attach(plot_data)
    
    for (im in 1:34) {
      Region_AV1451_v1 = plot_data$area_label[im]
      funtion_glm = as.formula(paste0(cognition_domains,"~",Region_AV1451_v1,"+Diagnosis_N+",Region_AV1451_v1,":Diagnosis_N+amyloid_age+Sex+Edu+eTIV"))
      results = MASS::rlm(funtion_glm,data =stats_freesurfer_PRS)
  
      coef_table <- summary(results)$coefficients
      results_df <- as.data.frame(coef_table)
      colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
      results_df <- results_df %>% mutate(p_value = 2 * (1 - pnorm(abs(t_value))))
      p_values = sfsmisc::f.robftest(results,var = names(results$coefficients)[8])

      plot_data$interaction_Effect_case_ref[im] = results_df[8,1]
      plot_data$interaction_Std_case_ref[im] = results_df[8,2]
      plot_data$interaction_pvalues_case_ref[im] = p_values$p.value
    }
    
    for (im in 35:41) {
      Region_AV1451_v1 = plot_data$area_label[im]
      Region_AV1451_ALL = plot_data$area_label
      funtion_glm = as.formula(paste0(cognition_domains,"~",Region_AV1451_v1,"+Diagnosis_N+",Region_AV1451_v1,
                                      ":Diagnosis_N+amyloid_age+Sex+Edu+eTIV"))
      results = MASS::rlm(funtion_glm,data =stats_freesurfer_PRS)
      
      coef_table <- summary(results)$coefficients
      results_df <- as.data.frame(coef_table)
      colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
      results_df <- results_df %>% mutate(p_value = 2 * (1 - pnorm(abs(t_value))))
      p_values = sfsmisc::f.robftest(results,var = names(results$coefficients)[8])
      
      plot_data$interaction_Effect_case_ref[im] = results_df[8,1]
      plot_data$interaction_Std_case_ref[im] = results_df[8,2]
      plot_data$interaction_pvalues_case_ref[im] = p_values$p.value
      
    }
    plot_data <- plot_data %>% 
      dplyr::select(region,area_label,interaction_Effect_case_ref,interaction_Std_case_ref,interaction_pvalues_case_ref) %>%
      distinct(area_label,.keep_all = TRUE) %>%
      mutate(cognition = cognition_domains)
    
    return(plot_data)
}


get_interaction_model_regions <- function(measure = c("volume", "thickness")) {
  measure <- match.arg(measure)

  cortical_regions <- c(
    "bankssts","caudal anterior cingulate","caudal middle frontal",
    "cuneus","entorhinal","frontal pole","fusiform","inferior parietal",
    "inferior temporal","insula","isthmus cingulate","lateral occipital",
    "lateral orbitofrontal","lingual","medial orbitofrontal","middle temporal",
    "paracentral","parahippocampal","pars opercularis","pars orbitalis",
    "pars triangularis","pericalcarine","postcentral","posterior cingulate",
    "precentral","precuneus","rostral anterior cingulate","rostral middle frontal",
    "superior frontal","superior parietal","superior temporal","supramarginal",
    "temporal pole","transverse temporal"
  )

  subcortical_region_labels <- c(
    "thalamus proper","caudate","putamen","pallidum","hippocampus","amygdala","accumbens area"
  )
  subcortical_volume_vars <- c(
    "Thalamus_volume_mean","Caudate_volume_mean","Putamen_volume_mean",
    "Pallidum_volume_mean","Hippocampus_volume_mean","Amygdala_volume_mean",
    "Accumbens_area_volume_mean"
  )

  tibble(
    region = c(cortical_regions, subcortical_region_labels),
    area_label = c(
      str_c(str_replace_all(cortical_regions, " ", ""), "_", measure, "_mean"),
      subcortical_volume_vars
    )
  )
}

tidy_vif_output <- function(vif_vals) {
  if (is.matrix(vif_vals) || is.data.frame(vif_vals)) {
    vif_df <- as.data.frame(vif_vals)
    vif_col <- if ("GVIF^(1/(2*Df))" %in% colnames(vif_df)) {
      "GVIF^(1/(2*Df))"
    } else {
      colnames(vif_df)[1]
    }

    tibble(
      term = rownames(vif_df),
      vif = as.numeric(vif_df[[vif_col]])
    )
  } else {
    tibble(
      term = names(vif_vals),
      vif = as.numeric(vif_vals)
    )
  }
}

calculate_interaction_model_vif <- function(data, cognition_domain, measure = c("volume", "thickness")) {
  measure <- match.arg(measure)
  model_regions <- get_interaction_model_regions(measure)

  purrr::map_dfr(seq_len(nrow(model_regions)), function(i) {
    region_name <- model_regions$region[i]
    region_var <- model_regions$area_label[i]

    formula_vif <- as.formula(
      paste0(
        cognition_domain, " ~ ", region_var,
        " * Diagnosis_N + amyloid_age + Sex + Edu + eTIV"
      )
    )

    lm_fit <- lm(formula_vif, data = data)
    vif_tbl <- tidy_vif_output(suppressWarnings(car::vif(lm_fit)))

    vif_tbl %>%
      mutate(
        cognition = cognition_domain,
        region = region_name,
        area_label = region_var,
        measure = measure,
        .before = 1
      )
  })
}

summarise_interaction_model_vif <- function(data, cognition_domains, measure = c("volume", "thickness")) {
  measure <- match.arg(measure)

  vif_terms <- purrr::map_dfr(cognition_domains, function(domain) {
    calculate_interaction_model_vif(data, domain, measure = measure)
  })

  vif_summary <- vif_terms %>%
    group_by(measure, cognition, region, area_label) %>%
    summarise(
      max_vif_term = term[which.max(vif)],
      max_vif = max(vif, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    arrange(cognition, desc(max_vif), region)

  list(
    summary = vif_summary,
    terms = vif_terms
  )
}



make_plots_interaction_case_control <- function(re_AD){
  
  re_AD$area_label = str_split(re_AD$area_label,"_",simplify = TRUE)[,1]
  re_AD$region[re_AD$region == "choroid.plexus"] = "choroid plexus"
  DKT_lobes_mapping <- 
    read.csv("/Users/jiangxiaofan/Desktop/projects/project_ADC_mega_atrophy_paper/Chenyang_ADC_paper/DKT_lobes_mapping_plot.csv")
  
  DKT_lobes_mapping$region[DKT_lobes_mapping$region == "choroid.plexus"] = "choroid plexus"
  
  re_AD <- merge(re_AD,DKT_lobes_mapping[,c("region","lobes")],by.x = "region",all.x = TRUE) %>% 
    filter(!region == "choroid plexus")

  re_AD$region <- factor(re_AD$region,levels = unique(DKT_lobes_mapping$region))
  re_AD$lobes <- factor(re_AD$lobes,
                        levels = c("Frontal Lobe","Temporal Lobe","Occipital Lobe","Parietal Lobe","Insula and Cingulate","Subcortical"))
  
  plot_AD <- ggplot(re_AD, aes(region, cognition, col = interaction_Effect_case_ref)) +  
    geom_tile(color = "#b9b6b6", fill = "white", size = 1) + 
    geom_point(aes(size = abs(interaction_Effect_case_ref)), shape = 15) +  
    geom_point(aes(shape = interaction_pvalues_case_ref < .05), size = 2, color = "black") +
    scale_shape_manual(
      values = c(`FALSE` = NA, `TRUE` = 8),
      guide  = "none") +
    labs(x = NULL, y = NULL, color = NULL) +  
    scale_color_gradientn(colours=c("#7CCEDE","white","#F3E66C","#EB482F"),values = scales::rescale(c(-0.5,0,0.5,1)),limits = c(-0.5,1),oob  = scales::squish)+
    scale_x_discrete(expand = c(0, 0), position = "top") +
    scale_y_discrete(expand = c(0, 0)) +  
    scale_size(range = c(1, 10), guide = NULL) +  
    guides(color = guide_colorbar(direction = "vertical", reverse = FALSE,                                 
                                  barwidth = unit(1, "cm"),                                
                                  barheight = unit(15, "cm"))) +
    theme_common() +
    facet_grid2(.~lobes, scales = "free_x", space = "free_x",
                strip = strip_themed(background_x = elem_list_rect(fill = c("#617b36","#2a5371","#2b9a17","#fa7351","#886c90","#cb1315","grey")),
                text_x = elem_list_text(color = "white", size = 12)))
   
    return(plot_AD)
}



make_plots_interaction_each_domain <- function(re_AD){
  
  re_AD$area_label = str_split(re_AD$area_label,"_",simplify = TRUE)[,1]
  re_AD$region[re_AD$region == "choroid.plexus"] = "choroid plexus"
  DKT_lobes_mapping <- 
    read.csv("/Users/jiangxiaofan/Desktop/projects/project_ADC_mega_atrophy_paper/Chenyang_ADC_paper/DKT_lobes_mapping_plot.csv")
  
  DKT_lobes_mapping$region[DKT_lobes_mapping$region == "choroid.plexus"] = "choroid plexus"
  re_AD <- merge(re_AD,DKT_lobes_mapping[,c("region","lobes")],by.x = "region",all.x = TRUE)
  re_AD$region <- factor(re_AD$region,levels = unique(DKT_lobes_mapping$region))
  re_AD$lobes <- factor(re_AD$lobes,
                        levels = c("Frontal Lobe","Temporal Lobe","Occipital Lobe","Parietal Lobe","Insula and Cingulate","Subcortical","ventricles"))
  
  plot_AD <- ggplot(re_AD, aes(region, cognition, col = interaction_Effect_case_ref)) +  
    geom_tile(color = "#b9b6b6", fill = "white", size = 1) + 
    geom_point(aes(size = abs(interaction_Effect_case_ref)), shape = 15) +  
    geom_point(aes(shape = interaction_pvalues_case_ref < .05), size = 2, color = "black") +
    scale_shape_manual(
      values = c(`FALSE` = NA, `TRUE` = 8),
      guide  = "none") +
    labs(x = NULL, y = NULL, color = NULL) +  
    scale_color_gradientn(colours=c("#7CCEDE","white","#F3E66C","#EB482F"),values = scales::rescale(c(-1.5,0,0.3,1,2)),limits = c(-1.5,2))+
    scale_x_discrete(expand = c(0, 0), position = "top") +
    scale_y_discrete(expand = c(0, 0)) +  
    scale_size(range = c(1, 10), guide = NULL) +  
    guides(color = guide_colorbar(direction = "vertical", reverse = FALSE,                                 
                                  barwidth = unit(0.5, "cm"),                                
                                  barheight = unit(15, "cm"))) +
    theme_common() +
    facet_grid2(.~lobes, scales = "free_x", space = "free_x",
                strip = strip_themed(background_x = elem_list_rect(fill = c("#617b36","#2a5371","#2b9a17","#fa7351","#886c90","#cb1315","grey")),text_x = elem_list_text(color = "white", size = 12)))
   
    return(plot_AD)
}


analysis_glm_cognition_mean <- function(stats_freesurfer_PRS,cognition_domains){
  
    list=c("bankssts","caudal anterior cingulate","caudal middle frontal",
        "cuneus","entorhinal","frontal pole","fusiform","inferior parietal",
        "inferior temporal","insula","isthmus cingulate","lateral occipital",
        "lateral orbitofrontal","lingual","medial orbitofrontal","middle temporal",
        "paracentral","parahippocampal","pars opercularis","pars orbitalis",
        "pars triangularis","pericalcarine","postcentral","posterior cingulate",
        "precentral","precuneus","rostral anterior cingulate","rostral middle frontal",
        "superior frontal","superior parietal","superior temporal","supramarginal",
        "temporal pole","transverse temporal","thalamus proper","caudate","putamen",
        "pallidum","hippocampus","amygdala","accumbens area","choroid.plexus")
    
    list_sub = c("Thalamus_volume_mean","Caudate_volume_mean","Putamen_volume_mean",
                 "Pallidum_volume_mean","Hippocampus_volume_mean","Amygdala_volume_mean",
                 "Accumbens_area_volume_mean","choroid_plexus_volume_mean")
    CT_names <- c(str_c(str_replace_all(list[1:34]," ",""),"_volume_mean"),list_sub)
    
    DTK_Labes_For_Plot = list
    plot_data_L = as.data.frame(DTK_Labes_For_Plot)
    colnames(plot_data_L)[1] = "region"
    plot_data = plot_data_L
    plot_data$area_label = CT_names
    plot_data$Effect=NA
    plot_data$P_value=NA
    attach(plot_data)
    
    for (im in 1:34) {
      Region_AV1451_v1 = plot_data$area_label[im]
      funtion_glm = paste(cognition_domains,"~",Region_AV1451_v1,"+amyloid_age+Sex+Edu+eTIV")
      results = glm(funtion_glm,family = gaussian,data =stats_freesurfer_PRS)
      x = summary(results)
      P_values_GLM = x$coefficients[2,3:4]
      plot_data$Effect[im] = x$coefficients[2,1]
      plot_data$P_value[im] = x$coefficients[2,4]
    }
    
    for (im in 35:42) {
      Region_AV1451_v1 = plot_data$area_label[im]
      funtion_glm = paste(cognition_domains,"~",Region_AV1451_v1,"+amyloid_age+Sex+Edu+eTIV")
      results = glm(funtion_glm,family = gaussian,data =stats_freesurfer_PRS)
      x = summary(results)
      P_values_GLM = x$coefficients[2,3:4]
      plot_data$Effect[im] = x$coefficients[2,1]
      plot_data$P_value[im] = x$coefficients[2,4]
    }
    return(plot_data)
}


theme_pub <- function(base_size = 20) {
  theme_classic(base_size = base_size) +
    theme(
      legend.position = "none",
      legend.title = element_text(size = base_size + 2, face = "bold"),
      legend.text = element_text(size = base_size),
      axis.title.x = element_text(size = base_size + 4, face = "bold", margin = margin(t = 10)),
      axis.title.y = element_text(size = base_size + 4, face = "bold", margin = margin(r = 10)),
      axis.text.x = element_text(size = base_size + 2),
      axis.text.y = element_text(size = base_size + 2),
      panel.grid.major.y = element_line(color = "gray90"),
      panel.grid.major.x = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      axis.line = element_line(color = "black")
    )
}

```

# Cortical thickness vs. grey matter volume: within-region scatter plots

```{r, fig.width=18, fig.height=22}
cortical_regions <- c(
  "bankssts","caudalanteriorcingulate","caudalmiddlefrontal",
  "cuneus","entorhinal","frontalpole","fusiform","inferiorparietal",
  "inferiortemporal","insula","isthmuscingulate","lateraloccipital",
  "lateralorbitofrontal","lingual","medialorbitofrontal","middletemporal",
  "paracentral","parahippocampal","parsopercularis","parsorbitalis",
  "parstriangularis","pericalcarine","postcentral","posteriorcingulate",
  "precentral","precuneus","rostralanteriorcingulate","rostralmiddlefrontal",
  "superiorfrontal","superiorparietal","superiortemporal","supramarginal",
  "temporalpole","transversetemporal"
)

# Build long data: one row per subject × region
scatter_data <- map_dfr(cortical_regions, function(r) {
  vol_col <- paste0(r, "_volume_mean")
  thi_col <- paste0(r, "_thickness_mean")
  if (!all(c(vol_col, thi_col) %in% names(data_ADC_analysis_cog))) return(NULL)
  data_ADC_analysis_cog %>%
    dplyr::select(vol = all_of(vol_col), thi = all_of(thi_col)) %>%
    drop_na() %>%
    mutate(region = r)
})

# Per-region Pearson r + FDR for panel labels
cor_stats <- scatter_data %>%
  group_by(region) %>%
  summarise(
    r       = cor(vol, thi, use = "complete.obs"),
    p_value = cor.test(vol, thi)$p.value,
    .groups = "drop"
  ) %>%
  mutate(fdr = p.adjust(p_value, "fdr"),
         label = sprintf("r = %.2f\nFDR = %.3f", r, fdr))

# Order panels by descending r
region_order <- cor_stats %>% arrange(desc(r)) %>% pull(region)

scatter_data <- scatter_data %>%
  mutate(region = factor(region, levels = region_order)) %>%
  left_join(cor_stats %>% dplyr::select(region, r), by = "region")

# Pretty region labels for facet strips
region_labels <- setNames(
  gsub("([a-z])([A-Z])", "\\1 \\2",
       str_to_title(gsub("([a-z])([A-Z])", "\\1 \\2", region_order))),
  region_order
)

ggplot(scatter_data, aes(x = vol, y = thi)) +
  geom_point(alpha = 0.25, size = 0.8, colour = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, colour = "#D32F2F",
              linewidth = 0.8, fill = "#FFCDD2") +
  geom_text(
    data = cor_stats %>% mutate(region = factor(region, levels = region_order)),
    aes(label = label),
    x = -Inf, y = Inf, hjust = -0.08, vjust = 1.2,
    size = 2.8, inherit.aes = FALSE
  ) +
  facet_wrap(~ region, scales = "free", ncol = 6,
             labeller = labeller(region = region_labels)) +
  labs(
    title    = "Within-region correlation: grey matter volume vs. cortical thickness",
    subtitle = "Panels ordered by Pearson r (high → low); regression line with 95% CI",
    x        = "Grey matter volume (mean)",
    y        = "Cortical thickness (mean)"
  ) +
  theme_bw(base_size = 10) +
  theme(
    strip.background = element_rect(fill = "grey92", colour = NA),
    strip.text       = element_text(size = 8, face = "bold"),
    axis.text        = element_text(size = 7),
    panel.grid.minor = element_blank()
  )

ggsave("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure7/volume_thickness_scatterplots.pdf",
       width = 18, height = 22, units = "in", dpi = 300)  
```



# FTD: interaction effects - robust linear regression model
```{r,include=FALSE}

data_N_A_analysis_FTD <- data_ADC_analysis_cog %>%
  filter(D_diagnosis %in% c("FTD")) %>%
  mutate(Diagnosis_N = factor(Diagnosis_N,c("FTD(A-)","FTD(A+)"))) %>%
  filter(!I_ID == 3329)
  
re_EpisodicMemory <- analysis_glm_cognition_case_control_interaction_thickness_robust(data_N_A_analysis_FTD,"n_memory_z")
re_ExecutiveFunctioning <- analysis_glm_cognition_case_control_interaction_thickness_robust(data_N_A_analysis_FTD,"n_executive_z")
re_Attention <- analysis_glm_cognition_case_control_interaction_thickness_robust(data_N_A_analysis_FTD,"n_attsp_z") 
re_Language <- analysis_glm_cognition_case_control_interaction_thickness_robust(data_N_A_analysis_FTD,"n_language_z") 
re_visualspatial <- analysis_glm_cognition_case_control_interaction_thickness_robust(data_N_A_analysis_FTD,"n_visualspatial_z") 

re_EpisodicMemory <- re_EpisodicMemory %>% 
  dplyr::select(region,area_label,interaction_Effect_case_ref,interaction_Std_case_ref,interaction_pvalues_case_ref) %>%
  distinct(area_label,.keep_all = TRUE) %>%
  mutate(cognition = "Episodic Memory")
  
re_ExecutiveFunctioning <- re_ExecutiveFunctioning %>% 
  dplyr::select(region,area_label,interaction_Effect_case_ref,interaction_Std_case_ref,interaction_pvalues_case_ref) %>%
  mutate(cognition = "Executive Functioning") %>%
  distinct(area_label,.keep_all = TRUE)

re_Attention <- re_Attention %>% 
  dplyr::select(region,area_label,interaction_Effect_case_ref,interaction_Std_case_ref,interaction_pvalues_case_ref) %>%
  mutate(cognition = "Attention") %>%
  distinct(area_label,.keep_all = TRUE)

re_Language <- re_Language %>% 
  dplyr::select(region,area_label,interaction_Effect_case_ref,interaction_Std_case_ref,interaction_pvalues_case_ref) %>%
  mutate(cognition = "Language") %>%
  distinct(area_label,.keep_all = TRUE)

re_visualspatial <- re_visualspatial %>% 
  dplyr::select(region,area_label,interaction_Effect_case_ref,interaction_Std_case_ref,interaction_pvalues_case_ref) %>%
  mutate(cognition = "Visual Spatial") %>%
  distinct(area_label,.keep_all = TRUE)

re_AD <- bind_rows(re_EpisodicMemory,re_ExecutiveFunctioning,re_Attention,re_Language)
re_AD$cognition = factor(re_AD$cognition,levels = c("Language","Attention","Executive Functioning","Episodic Memory"))
plot_FTD <- make_plots_interaction_case_control(re_AD)

ggsave("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure7/Figure2_FTD_hemapmap.pdf",plot_FTD,width = 17,height = 5,units = "in",dpi = 300)
#ggsave("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/plots/plots_association/robust/Figure2_FTD_hemapmap_LEGEND.pdf",plot_FTD,width = 17,height = 10,units = "in",dpi = 300)

re_AD <- re_AD %>% 
  mutate(interaction_fdr_case_ref = p.adjust(interaction_pvalues_case_ref,"fdr")) %>%
  mutate(across(where(is.numeric), ~ signif(., 2))) 

writexl::write_xlsx(re_AD,"/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure7/interactions_FTD.xlsx")
```


# create FTD scatter TABLE
```{r}
re_AD_2 <- re_AD %>% 
  mutate(cognition = case_when(
    cognition == "Language" ~ "n_language_z",
    cognition == "Episodic Memory" ~ "n_memory_z",
    cognition == "Executive Functioning" ~ "n_executive_z",
    cognition == "Attention" ~ "n_attsp_z",
    TRUE ~ NA_character_))
  #filter(!is.na(cognition))

df_ALL <- as.data.frame(matrix(NA, nrow =nrow(re_AD_2), ncol = 8)) 
names(df_ALL) = c("MRI","Cog","beta","std","pvalue","beta_AD","std_AD","pvalue_AD")

for (im in 1:nrow(re_AD_2)) {

covars_txt <- if (grepl("thickness_mean", re_AD_2$area_label[im])) {
  "+amyloid_age+Sex+Edu"
} else {
  "+amyloid_age+Sex+Edu+eTIV"
}

data_N_A_analysis_tmp <- data_N_A_analysis_FTD %>% filter(Diagnosis_N == "FTD(A-)")

funtion_glm = as.formula(str_c(re_AD_2$cognition[im],"~", re_AD_2$area_label[im],covars_txt))
results = MASS::rlm(funtion_glm,data =data_N_A_analysis_tmp)
coef_table <- summary(results)$coefficients
results_df <- as.data.frame(coef_table)
colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
results_df <- results_df %>% mutate(p_value = 2 * (1 - pnorm(abs(t_value))))
coef_info <- results_df[re_AD_2$area_label[im], ]
beta_txt_minus <- signif(coef_info$Estimate,2)
std_txt_minus <- signif(coef_info$Std_Error,2)

p_values = sfsmisc::f.robftest(results,var = names(results$coefficients)[2])
pval_txt_minus <- signif(p_values$p.value,2)

df_ALL[im,1] = re_AD_2$area_label[im]
df_ALL[im,2] = re_AD_2$cognition[im]

df_ALL[im,3] = beta_txt_minus
df_ALL[im,4] = std_txt_minus
df_ALL[im,5] = pval_txt_minus

data_N_A_analysis_tmp <- data_N_A_analysis_FTD %>% filter(Diagnosis_N == "FTD(A+)")

funtion_glm = as.formula(str_c(re_AD_2$cognition[im],"~", re_AD_2$area_label[im],"+amyloid_age+Sex+Edu+eTIV"))

funtion_glm = as.formula(str_c(re_AD_2$cognition[im],"~", re_AD_2$area_label[im],covars_txt))
results = MASS::rlm(funtion_glm,data =data_N_A_analysis_tmp)
coef_table <- summary(results)$coefficients
results_df <- as.data.frame(coef_table)
colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
results_df <- results_df %>% mutate(p_value = 2 * (1 - pnorm(abs(t_value))))
coef_info <- results_df[re_AD_2$area_label[im], ]
beta_txt_minus <- signif(coef_info$Estimate,2)
std_txt_minus <- signif(coef_info$Std_Error,2)

p_values = sfsmisc::f.robftest(results,var = names(results$coefficients)[2])
pval_txt_minus <- signif(p_values$p.value,2)

df_ALL[im,6] = beta_txt_minus
df_ALL[im,7] = std_txt_minus
df_ALL[im,8] = pval_txt_minus

}

writexl::write_xlsx(df_ALL,"/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure7/test_FTD_plus.xlsx")
```



# FTD scatter plots
```{r}
data_N_A_analysis_tmp <- data_N_A_analysis_FTD %>% filter(Diagnosis_N == "FTD(A+)")
funtion_glm = as.formula("n_language_z~Putamen_volume_mean+amyloid_age+Sex+Edu+eTIV")
results = MASS::rlm(funtion_glm,data =data_N_A_analysis_tmp)
coef_table <- summary(results)$coefficients
results_df <- as.data.frame(coef_table)
colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
results_df <- results_df %>% mutate(p_value = 2 * (1 - pnorm(abs(t_value))))
coef_info <- results_df["Putamen_volume_mean", ]
beta_txt <- sprintf("BetaA+ = %.3f", coef_info["Estimate"])
p_values = sfsmisc::f.robftest(results,var = names(results$coefficients)[2])
pval_txt <- ifelse(p_values$p.value < 0.001, "p < 0.001", sprintf("p = %.3f", p_values$p.value))

plot_FTD_1 <- ggplot(data_N_A_analysis_FTD, aes(x = Putamen_volume_mean, 
                                              y = n_language_z,
                                              color = Diagnosis_N,
                                              linetype = Diagnosis_N)) +
  geom_point(alpha = 0.6, size = 4) +
  geom_smooth(method = "lm", se = TRUE, aes(fill = Diagnosis_N)) +
  labs(x = "Putamen volume", y = NULL) +
  labs(y = "Language domain", y = NULL) +
  scale_x_continuous(limits = c(-2, 3), breaks = seq(-2, 3, by = 1)) +
  scale_y_continuous(limits = c(-4, 1)) +
  scale_linetype_manual(values = c("FTD(A-)" = "solid", "FTD(A+)" = "dashed")) +
  scale_color_manual(values = c("FTD(A-)" = "#507991", "FTD(A+)" = "#D9829D")) +
  scale_fill_manual(values = c("FTD(A-)" = "#507991", "FTD(A+)" = "#D9829D")) +
  annotate("text",x = 1, y = -2.6,label = expression("A-:" ~ beta == 0.089 * "," ~ italic(p) == 0.15),size = 7,hjust = 0) +
  annotate("text",x = 1, y = -3.4,label = expression("A+:" ~ beta == 0.44 * "," ~ italic(p) == 0.0062),size = 7,hjust = 0) +
  theme_pub()

################################################  Figure2_FTD_scatter2.pdf
data_N_A_analysis_tmp <- data_N_A_analysis_FTD %>% filter(Diagnosis_N == "FTD(A-)")
funtion_glm = as.formula("n_attsp_z~posteriorcingulate_thickness_mean+amyloid_age+Sex+Edu+eTIV")
results = MASS::rlm(funtion_glm,data =data_N_A_analysis_tmp)
coef_table <- summary(results)$coefficients
results_df <- as.data.frame(coef_table)
colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
results_df <- results_df %>% mutate(p_value = 2 * (1 - pnorm(abs(t_value))))

coef_info <- results_df["posteriorcingulate_thickness_mean", ]
beta_txt <- sprintf("Beta = %.3f", coef_info["Estimate"])
p_values = sfsmisc::f.robftest(results,var = names(results$coefficients)[2])
pval_txt <- ifelse(p_values$p.value < 0.001, "p < 0.001", sprintf("p = %.3f", p_values$p.value))

plot_FTD_2 <- ggplot(data_N_A_analysis_FTD, aes(x = posteriorcingulate_thickness_mean, 
                                              y = n_attsp_z,
                                              color = Diagnosis_N,
                                              linetype = Diagnosis_N)) +
  geom_point(alpha = 0.6, size = 4) +
  geom_smooth(method = "lm", se = TRUE, aes(fill = Diagnosis_N)) +
  labs(x = "Posterior cingulate thickness", y = NULL) +
  labs(y = "Attention domain", y = NULL) +
  scale_x_continuous(limits = c(-2, 3), breaks = seq(-2, 3, by = 1)) +
  scale_y_continuous(limits = c(-4, 1)) +
  scale_linetype_manual(values = c("FTD(A-)" = "solid", "FTD(A+)" = "dashed")) +
  scale_color_manual(values = c("FTD(A-)" = "#507991", "FTD(A+)" = "#D9829D")) +
  scale_fill_manual(values = c("FTD(A-)" = "#507991", "FTD(A+)" = "#D9829D")) +
  annotate("text",x = 1, y = -2.6,label = expression("A-:" ~ beta == -0.028 * "," ~ italic(p) == 0.69),size = 7,hjust = 0) +
  annotate("text",x = 1, y = -3.4,label = expression("A+:" ~ beta == 0.43 * "," ~ italic(p) == 0.0016),size = 7,hjust = 0) +
  theme_pub()

ggsave("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure7/Figure2_FTD_scatter_ALL_1.pdf",
  plot_FTD_1,width = 17,height = 5,units = "in",dpi = 300)
ggsave("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure7/Figure2_FTD_scatter_ALL_2.pdf",
  plot_FTD_2,width = 17,height = 5,units = "in",dpi = 300)
```



# DLB: interaction effects - robust linear regression model
```{r}
data_N_A_analysis_DLB <- data_ADC_analysis_cog %>%
  filter(D_diagnosis %in% c("DLB")) %>%
  mutate(Diagnosis_N = factor(Diagnosis_N,c("DLB(A-)","DLB(A+)")))

re_EpisodicMemory <- analysis_glm_cognition_case_control_interaction_thickness_robust(data_N_A_analysis_DLB,"n_memory_z")
re_ExecutiveFunctioning <- analysis_glm_cognition_case_control_interaction_thickness_robust(data_N_A_analysis_DLB,"n_executive_z")
re_Attention <- analysis_glm_cognition_case_control_interaction_thickness_robust(data_N_A_analysis_DLB,"n_attsp_z") 
re_Language <- analysis_glm_cognition_case_control_interaction_thickness_robust(data_N_A_analysis_DLB,"n_language_z") 
re_visualspatial <- analysis_glm_cognition_case_control_interaction_thickness_robust(data_N_A_analysis_DLB,"n_visualspatial_z") 

re_EpisodicMemory <- re_EpisodicMemory %>% 
  dplyr::select(region,area_label,interaction_Effect_case_ref,interaction_Std_case_ref,interaction_pvalues_case_ref) %>%
  distinct(area_label,.keep_all = TRUE) %>%
  mutate(cognition = "Episodic Memory")
  
re_ExecutiveFunctioning <- re_ExecutiveFunctioning %>% 
  dplyr::select(region,area_label,interaction_Effect_case_ref,interaction_Std_case_ref,interaction_pvalues_case_ref) %>%
  mutate(cognition = "Executive Functioning") %>%
  distinct(area_label,.keep_all = TRUE)

re_Attention <- re_Attention %>% 
  dplyr::select(region,area_label,interaction_Effect_case_ref,interaction_Std_case_ref,interaction_pvalues_case_ref) %>%
  mutate(cognition = "Attention") %>%
  distinct(area_label,.keep_all = TRUE)

re_Language <- re_Language %>% 
  dplyr::select(region,area_label,interaction_Effect_case_ref,interaction_Std_case_ref,interaction_pvalues_case_ref) %>%
  mutate(cognition = "Language") %>%
  distinct(area_label,.keep_all = TRUE)

re_visualspatial <- re_visualspatial %>% 
  dplyr::select(region,area_label,interaction_Effect_case_ref,interaction_Std_case_ref,interaction_pvalues_case_ref) %>%
  mutate(cognition = "Visual Spatial") %>%
  distinct(area_label,.keep_all = TRUE)

re_AD <- bind_rows(re_EpisodicMemory,re_ExecutiveFunctioning,re_Attention,re_Language)
re_AD$cognition = factor(re_AD$cognition,levels = c("Language","Attention","Executive Functioning","Episodic Memory"))
plot_DLB <- make_plots_interaction_case_control(re_AD)

ggsave("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure7_8/Figure2_DLB_hemapmap.pdf",plot_DLB,width = 17,height = 5,units = "in",dpi = 300)

re_AD <- re_AD %>% 
  mutate(interaction_fdr_case_ref = p.adjust(interaction_pvalues_case_ref,"fdr")) %>%
  mutate(across(where(is.numeric), ~ signif(., 2)))
writexl::write_xlsx(re_AD,"/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure7/interactions_DLB.xlsx")

```


# VIF check for interaction models
```{r}
thickness_domains <- c("n_memory_z", "n_executive_z", "n_attsp_z", "n_language_z", "n_visualspatial_z")

VIF_FTD <- summarise_interaction_model_vif(
  data = data_N_A_analysis_FTD,
  cognition_domains = thickness_domains,
  measure = "thickness"
)

VIF_DLB <- summarise_interaction_model_vif(
  data = data_N_A_analysis_DLB,
  cognition_domains = thickness_domains,
  measure = "thickness"
)

VIF_model_all <- bind_rows(
  VIF_FTD$summary %>% mutate(group = "FTD"),
  VIF_DLB$summary %>% mutate(group = "DLB")
) %>%
  mutate(max_vif = round(max_vif, 3)) %>%
  select(group, everything())

VIF_terms_all <- bind_rows(
  VIF_FTD$terms %>% mutate(group = "FTD"),
  VIF_DLB$terms %>% mutate(group = "DLB")
) %>%
  mutate(vif = round(vif, 3)) %>%
  select(group, everything())

VIF_summary_all <- bind_rows(
  VIF_model_all %>%
    group_by(group, measure) %>%
    summarise(
      n_models = n(),
      median_max_vif = round(median(max_vif, na.rm = TRUE), 3),
      q1_max_vif = round(quantile(max_vif, probs = 0.25, na.rm = TRUE), 3),
      q3_max_vif = round(quantile(max_vif, probs = 0.75, na.rm = TRUE), 3),
      min_max_vif = round(min(max_vif, na.rm = TRUE), 3),
      max_max_vif = round(max(max_vif, na.rm = TRUE), 3),
      .groups = "drop"
    ),
  VIF_model_all %>%
    summarise(
      group = "All",
      measure = dplyr::first(measure),
      n_models = n(),
      median_max_vif = round(median(max_vif, na.rm = TRUE), 3),
      q1_max_vif = round(quantile(max_vif, probs = 0.25, na.rm = TRUE), 3),
      q3_max_vif = round(quantile(max_vif, probs = 0.75, na.rm = TRUE), 3),
      min_max_vif = round(min(max_vif, na.rm = TRUE), 3),
      max_max_vif = round(max(max_vif, na.rm = TRUE), 3)
    )
) %>%
  arrange(factor(group, levels = c("FTD", "DLB", "All")))

VIF_overview <- VIF_model_all %>%
  group_by(group, measure, cognition) %>%
  summarise(
    n_models = n(),
    median_max_vif = round(median(max_vif, na.rm = TRUE), 3),
    max_max_vif = round(max(max_vif, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  arrange(group, cognition)

writexl::write_xlsx(
  list(
    summary = VIF_summary_all,
    model_level = VIF_model_all,
    term_level = VIF_terms_all,
    cognition_overview = VIF_overview
  ),
  "/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure7/Supplementary_Table_13_VIF_interaction_models_thickness.xlsx"
)
```


# create DLB scatter plots TABLE
```{r}
re_AD_2 <- re_AD %>% 
  mutate(cognition = case_when(
    cognition == "Language" ~ "n_language_z",
    cognition == "Episodic Memory" ~ "n_memory_z",
    cognition == "Executive Functioning" ~ "n_executive_z",
    cognition == "Attention" ~ "n_attsp_z",
    TRUE ~ NA_character_))

df_ALL <- as.data.frame(matrix(NA, nrow =nrow(re_AD_2), ncol = 8)) 
names(df_ALL) = c("MRI","Cog","beta","std","pvalue","beta_AD","std_AD","pvalue_AD")

for (im in 1:nrow(re_AD_2)) {
  
data_N_A_analysis_tmp <- data_N_A_analysis_DLB %>% filter(Diagnosis_N == "DLB(A-)")

funtion_glm = as.formula(str_c(re_AD_2$cognition[im],"~", re_AD_2$area_label[im],"+amyloid_age+Sex+Edu+eTIV"))
results = MASS::rlm(funtion_glm,data =data_N_A_analysis_tmp)
coef_table <- summary(results)$coefficients
results_df <- as.data.frame(coef_table)
colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
results_df <- results_df %>% mutate(p_value = 2 * (1 - pnorm(abs(t_value))))
coef_info <- results_df[re_AD_2$area_label[im], ]
beta_txt_minus <- signif(coef_info$Estimate,2)
std_txt_minus <- signif(coef_info$Std_Error,2)

p_values = sfsmisc::f.robftest(results,var = names(results$coefficients)[2])
pval_txt_minus <- signif(p_values$p.value,2)

df_ALL[im,1] = re_AD_2$area_label[im]
df_ALL[im,2] = re_AD_2$cognition[im]

df_ALL[im,3] = beta_txt_minus
df_ALL[im,4] = std_txt_minus
df_ALL[im,5] = pval_txt_minus


data_N_A_analysis_tmp <- data_N_A_analysis_DLB %>% filter(Diagnosis_N == "DLB(A+)")

funtion_glm = as.formula(str_c(re_AD_2$cognition[im],"~", re_AD_2$area_label[im],"+amyloid_age+Sex+Edu+eTIV"))
results = MASS::rlm(funtion_glm,data =data_N_A_analysis_tmp)
coef_table <- summary(results)$coefficients
results_df <- as.data.frame(coef_table)
colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
results_df <- results_df %>% mutate(p_value = 2 * (1 - pnorm(abs(t_value))))
coef_info <- results_df[re_AD_2$area_label[im], ]
beta_txt_minus <- signif(coef_info$Estimate,2)
std_txt_minus <- signif(coef_info$Std_Error,2)

p_values = sfsmisc::f.robftest(results,var = names(results$coefficients)[2])
pval_txt_minus <- signif(p_values$p.value,2)

df_ALL[im,6] = beta_txt_minus
df_ALL[im,7] = std_txt_minus
df_ALL[im,8] = pval_txt_minus

}

df_ALL <- df_ALL %>% 
  mutate(across(where(is.numeric), ~ signif(., 2)))

writexl::write_xlsx(df_ALL,"/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure7/test_DLB_plus.xlsx")
```



# DLB scatter plots
```{r}
######################## Figure2_FTD_scatter1.pdf
data_N_A_analysis_tmp <- data_N_A_analysis_DLB %>% filter(Diagnosis_N == "DLB(A-)")
funtion_glm = as.formula("n_language_z~Caudate_volume_mean+amyloid_age+Sex+Edu+eTIV")
results = MASS::rlm(funtion_glm,data =data_N_A_analysis_tmp)
coef_table <- summary(results)$coefficients
results_df <- as.data.frame(coef_table)
colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
results_df <- results_df %>% mutate(p_value = 2 * (1 - pnorm(abs(t_value))))
coef_info <- results_df["Caudate_volume_mean", ]
beta_txt <- sprintf("Beta = %.3f", coef_info["Estimate"])
p_values = sfsmisc::f.robftest(results,var = names(results$coefficients)[2])
pval_txt <- ifelse(p_values$p.value < 0.001, "p < 0.001", sprintf("p = %.3f", p_values$p.value))


plot_DLB_1 <- ggplot(data_N_A_analysis_DLB, aes(x = Caudate_volume_mean, 
                                              y = n_language_z,
                                              color = Diagnosis_N,
                                              linetype = Diagnosis_N)) +
  geom_point(alpha = 0.6, size = 4) +
  geom_smooth(method = "lm", se = TRUE, aes(fill = Diagnosis_N)) +
  labs(x = "Caudate volume", y = NULL) +
  labs(y = "Language domain", y = NULL) +
  scale_x_continuous(limits = c(-2, 3), breaks = seq(-2, 3, by = 1)) +
  scale_y_continuous(limits = c(-4, 1)) +
  scale_linetype_manual(values = c("DLB(A-)" = "solid", "DLB(A+)" = "dashed")) +
  scale_color_manual(values = c("DLB(A-)" = "#507991", "DLB(A+)" = "#D9829D")) +
  scale_fill_manual(values = c("DLB(A-)" = "#507991", "DLB(A+)" = "#D9829D")) +
  annotate("text",x = 1, y = -2.6,label = expression("A-:" ~ beta == -0.12 * "," ~ italic(p) == 0.087),size = 7,hjust = 0) +
  annotate("text",x = 1, y = -3.4,label = expression("A+:" ~ beta == 0.11 * "," ~ italic(p) == 0.12),size = 7,hjust = 0) +
  theme_pub()

######################################################## Figure3_FTD_scatter1.pdf
data_N_A_analysis_tmp <- data_N_A_analysis_DLB %>% filter(Diagnosis_N == "DLB(A-)")
funtion_glm = as.formula("n_language_z~inferiortemporal_thickness_mean+amyloid_age+Sex+Edu+eTIV")
results = MASS::rlm(funtion_glm,data =data_N_A_analysis_tmp)
coef_table <- summary(results)$coefficients
results_df <- as.data.frame(coef_table)
colnames(results_df) <- c("Estimate", "Std_Error", "t_value")
results_df <- results_df %>% mutate(p_value = 2 * (1 - pnorm(abs(t_value))))
coef_info <- results_df["inferiortemporal_thickness_mean", ]
beta_txt <- sprintf("Beta = %.3f", coef_info["Estimate"])
p_values = sfsmisc::f.robftest(results,var = names(results$coefficients)[2])
pval_txt <- ifelse(p_values$p.value < 0.001, "p < 0.001", sprintf("p = %.3f", p_values$p.value))

plot_DLB_2 <- ggplot(data_N_A_analysis_DLB, aes(x = inferiortemporal_thickness_mean, 
                                                y = n_language_z,
                                                color = Diagnosis_N,
                                                linetype = Diagnosis_N)) +
  geom_point(alpha = 0.6, size = 4) +
  geom_smooth(method = "lm", se = TRUE, aes(fill = Diagnosis_N)) +
  labs(x = "inferior temporal thickness", y = NULL) +
  labs(y = "Language domain", y = NULL) +
  scale_x_continuous(limits = c(-2, 3), breaks = seq(-2, 3, by = 1)) +
  scale_y_continuous(limits = c(-4, 1)) +
  scale_linetype_manual(values = c("DLB(A-)" = "solid", "DLB(A+)" = "dashed")) +
  scale_color_manual(values = c("DLB(A-)" = "#507991", "DLB(A+)" = "#D9829D")) +
  scale_fill_manual(values = c("DLB(A-)" = "#507991", "DLB(A+)" = "#D9829D")) +
  annotate("text",x = 1, y = -2.6,label = expression("A-:" ~ beta == 0.0053 * "," ~ italic(p) == 0.96),size = 7,hjust = 0) +
  annotate("text",x = 1, y = -3.4,label = expression("A+:" ~ beta == 0.34 * "," ~ italic(p) == 0.012),size = 7,hjust = 0) +
  theme_pub()


ggsave("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure7/Figure2_DLB_scatter_ALL_1.pdf",plot_DLB_1,width = 17,height = 5,units = "in",dpi = 300)
ggsave("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/documents/Rebuttal/results/S_Figure7/Figure2_DLB_scatter_ALL_2.pdf",plot_DLB_2,width = 17,height = 5,units = "in",dpi = 300)
```



## Global wise associations: AD, FTD and DLB
```{r,include=FALSE}
data_N_A_analysis_FTD <- data_ADC_analysis %>%
  filter(D_diagnosis %in% c("AD","FTD","DLB")) %>%
  mutate(D_diagnosis = factor(D_diagnosis,c("AD","FTD","DLB"))) %>%
  mutate(EpisodicMemory = n_memory_z,
         ExecutiveFunctioning = n_executive_z,
         Attention = n_attsp_z,
         Language = n_language_z)


re_EpisodicMemory <- analysis_glm_cognition_mean(data_N_A_analysis_FTD,"EpisodicMemory")
re_ExecutiveFunctioning <- analysis_glm_cognition_mean(data_N_A_analysis_FTD,"ExecutiveFunctioning")
re_Attention <- analysis_glm_cognition_mean(data_N_A_analysis_FTD,"Attention") 
re_Language <- analysis_glm_cognition_mean(data_N_A_analysis_FTD,"Language") 

```


# AD+FTD+DLB:
```{r,include=FALSE}
data_C1_domains <- data_ADC_analysis %>%
  filter(D_diagnosis %in% c("AD","FTD","DLB")) %>%
  mutate(D_diagnosis = factor(D_diagnosis,c("AD","FTD","DLB"))) %>%
  mutate(EpisodicMemory = n_memory_z,
         ExecutiveFunctioning = n_executive_z,
         Attention = n_attsp_z,
         Language = n_language_z)

theme_common <- function() {
  theme(
    axis.ticks.x = element_blank(),        
    axis.ticks.y = element_blank(),        
    panel.border = element_rect(fill = NA, color = "grey70", size = 2, linetype = "solid"),
    axis.text.x = element_text(size = 16, angle = 90, hjust = 0),
    axis.text.y = element_text(size = 16),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 16),
    legend.key.size = unit(1.5, "cm"),
    legend.box.spacing = unit(0.5, "cm"))
  
}

### sort as lobes
i = 3
DKT_lobes_mapping2 = read.csv("/Users/jiangxiaofan/Desktop/projects/project_ADC_mega_atrophy_paper/Chenyang_ADC_paper/DKT_lobes_mapping_2.csv")
group_case = c("AD","FTD","DLB")

data_analysis1_b <- data_C1_domains %>% filter(D_diagnosis %in% group_case[i])

re_1_AD <- data_analysis1_b %>% analysis_glm_cognition_mean(., "EpisodicMemory") 
re_2_AD <- data_analysis1_b %>% analysis_glm_cognition_mean(., "ExecutiveFunctioning")
re_3_AD <- data_analysis1_b %>% analysis_glm_cognition_mean(., "Attention")
re_4_AD <- data_analysis1_b %>% analysis_glm_cognition_mean(., "Language")

re_1_AD$cognition = "EpisodicMemory"
re_2_AD$cognition = "ExecutiveFunctioning"
re_3_AD$cognition = "Attention"
re_4_AD$cognition = "Language"


re_AD_excel <- bind_rows(re_4_AD,re_3_AD,re_2_AD,re_1_AD) %>% 
  mutate(p.adjust = p.adjust(P_value, "fdr", n = 210))
#writexl::write_xlsx(re_AD_excel,"/Users/jiangxiaofan/Desktop/projects/project_ADC_mega_atrophy_paper/Chenyang_ADC_paper/submission/figures/figure4/re_AD_excel.xlsx")

re_AD <- bind_rows(re_4_AD,re_3_AD,re_2_AD,re_1_AD) %>%
  #mutate(p.adjust = p.adjust(P_value, "fdr", n = 210)) %>%
  mutate(Effect = if_else(P_value > 0.05, NA_real_, Effect)) %>%
  mutate(cognition = factor(cognition,c("EpisodicMemory","ExecutiveFunctioning","Attention","Language")))


re_AD$area_label = str_split(re_AD$area_label,"_",simplify = TRUE)[,1]
re_AD$region[re_AD$region == "choroid.plexus"] = "choroid plexus"
DKT_lobes_mapping <- read.csv("/Users/jiangxiaofan/Desktop/projects/project_ADC_mega_atrophy_paper/Chenyang_ADC_paper/DKT_lobes_mapping_plot.csv")


DKT_lobes_mapping$region[DKT_lobes_mapping$region == "choroid.plexus"] = "choroid plexus"
re_AD <- merge(re_AD,DKT_lobes_mapping[,c("region","lobes")],by.x = "region",all.x = TRUE) %>%
  filter(!region == "choroid plexus")
re_AD$region <- factor(re_AD$region,levels = unique(DKT_lobes_mapping$region))
re_AD$lobes <- factor(re_AD$lobes,levels = c("Frontal Lobe","Temporal Lobe","Occipital Lobe","Parietal Lobe","Insula and Cingulate","Subcortical"))

plot_AD <- ggplot(re_AD, aes(region, cognition, col = Effect)) +  
  geom_tile(color = "#b9b6b6", fill = "white", size = 1) + 
  geom_point(aes(size = abs(Effect)), shape = 15) +
  #geom_point(aes(shape = P_value < .001), size = 2, color = "black") +
  #coord_equal() + 
  labs(x = NULL, y = NULL, color = NULL) +  
  scale_color_gradientn(colours=c("#7CCEDE","white","#F3E66C","#EB482F"),values = scales::rescale(c(-0.6,0,0.3,1.5)),limits = c(-0.8,1.5))+
  scale_x_discrete(expand = c(0, 0), position = "top") +
  scale_y_discrete(expand = c(0, 0)) +  
  scale_size(range = c(1, 10), guide = NULL) +  
  guides(color = guide_colorbar(direction = "vertical", reverse = FALSE,                                 
                                barwidth = unit(0.5, "cm"),                                
                                barheight = unit(15, "cm"))) +
  theme_common() +
  facet_grid2(.~lobes, scales = "free_x", space = "free_x",
              strip = strip_themed(background_x = elem_list_rect(fill = c("#617b36","#2a5371","#2b9a17","#fa7351","#886c90","#cb1315","grey")),text_x = elem_list_text(color = "white", size = 12))) +
  theme(legend.position = "none")
  

ggsave("/Users/jiangxiaofan/Desktop/projects/project_ADC_copathology/plots/plots_association/supp/Supp_Figure2_associations_pattern_DLB.pdf",plot_AD,width = 17,height = 5,units = "in",dpi = 300)

```
